---
title: "Final Report - R-Code - Exploring the possibility to predict student specialization choice from student profiles"
author: "Wessley Kameraad, René Freichel, Jan Simson"
date: "04.01.2021 - 29.01.2021"
output:
  html_document:
    toc: true
    toc_float: true
    toc_depth: 4
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
	echo = TRUE,
	message = FALSE,
	warning = FALSE
)
```

```{r, echo = FALSE, message = FALSE, warning = FALSE}
installed_packages <- installed.packages()[, "Package"]

## Install Missing Required Packages
if (!"tidyverse" %in% installed_packages) {
  install.packages("tidyverse")
  }
if (!"qgraph" %in% installed_packages) {
  install.packages("qgraph")
}
if (!"dplyr" %in% installed_packages) {
  install.packages("dplyr")
}
if (!"magrittr" %in% installed_packages) {
  install.packages("magrittr")
}
if (!"bootnet" %in% installed_packages) {
  install.packages("bootnet")
}
if (!"mgm" %in% installed_packages) {
  install.packages("mgm")
}
if (!"janitor" %in% installed_packages) {
  install.packages("janitor")
}

if (!"nnet" %in% installed_packages) {
  install.packages("nnet")
}
if (!requireNamespace("BiocManager", quietly = TRUE)){
  install.packages("BiocManager")
}
if (!"Rgraphviz" %in% installed_packages) {
  BiocManager::install("Rgraphviz")
}
    

## Load packages
library("tidyverse")
library("qgraph")
library("dplyr")
library("magrittr")
library("bootnet")
library("mgm")
library("janitor")
library("bnlearn")
library("Rgraphviz")
library("nnet")

# Set an initial seed for reproducibility
set.seed(2021)
```

```{r load_raw_data, message=FALSE, warning=FALSE}
cohorts <- c("1718", "1819", "1920")

# Load and combine the three CSVs
raw_data <- data.frame()
for (cohort in cohorts) {
  file <- paste0("data/anonymized_cohort", cohort, ".csv")
  
  cohort_data <- read_csv2(file) %>%
    clean_names() %>% 
    mutate(
      # Add a new column "cohort"
      cohort = cohort
    )

  
  # Add new cohorts data to joined data frame
  raw_data <- bind_rows(
    raw_data,
    cohort_data
  )
}
```

```{r tidy_data}
# Create a tidy dataframe with a subset of the data
full_tidy_data <- raw_data %>% 
  transmute(
    # Parse the grade as a decimal number
    grade = parse_number(
      offic_cijfer,
      locale = locale(decimal_mark = ",", grouping_mark = ".")
    ) %>%
      suppressWarnings,
    # Put all special grades (i.e. letters) into a separate column
    grade_special = if_else(is.na(grade), offic_cijfer, as.character(NA)),
    # Rename & select other columns
    specialisation = specialisatie,
    course_id = studiegids,
    student_id = masked_id,
    # The short description has different column names in different cohorts
    short_description = if_else(is.na(omschr_10), omschr_11, omschr_10),
    long_description = curs_omschr_l,
    cohort = cohort,
    rank = rangnummer,
    country = land
  )
```

```{r filter_data}
selected_courses <- c(
  "7201602PXY", # 1. Introduction Psychology & Cognition
  "7201604PXY", # 2. Research Methods & Statistics
  "7201605PXY", # 3. Developmental Psy
  "7201607PXY", # 4. Social & Work Psy
  "7201622PXY" # 5. Clinical Psychology & Brain
  # 5. is matched with "7201610PXY": Clinical & Biological Psy (1718)
)

# Merge the two courses across cohorts:
# Clinical Psychology & Brain (cohorts 1819, 1920)
# Clinical & Biological Psy (cohort 1718)
# and remove any instances where a student has taken both courses
# (that means they failed the first one, so that is dropped)
students_with_double_bio_courses <- full_tidy_data %>%
  filter(course_id %in% c("7201622PXY", "7201610PXY")) %>% 
  count(student_id) %>% 
  filter(n > 1) %>% 
  pull(student_id)
full_tidy_data <- full_tidy_data %>%
  filter(
    # Only keep the newer bio course if a student has both
    # (this means they failed to complete the old bio course)
    course_id != "7201610PXY" | !(student_id %in% students_with_double_bio_courses)
  ) %>% 
  mutate(
    course_id = if_else(course_id == "7201610PXY", "7201622PXY", course_id),
    short_description = if_else(
      course_id == "7201622PXY",
      "Clinical Psychology & Brain / Bio",
      short_description
    ),
    long_description = if_else(
      course_id == "7201622PXY", 
      "[Combined] Clinical Psychology & Brain + Clinical & Biological Psy", 
      long_description
    )
  )

data <- full_tidy_data %>%
  # Only include students who picked a specialisation
  filter(!is.na(specialisation) & specialisation != "0") %>% 
  filter(specialisation != "Algemene Psychologie") %>% 
  # Include only first year courses
  filter(course_id %in% selected_courses)
```

```{r}
course_abbrevations <- c(
  "7201602PXY" = "IPC",
  "7201604PXY" = "RMS",
  "7201605PXY" = "DEV",
  "7201607PXY" = "SWO",
  "7201622PXY" = "CB"
)

data_wide <- data %>%
  select(specialisation, grade, course_id, student_id) %>%
  mutate(course_id = recode(course_id, !!!course_abbrevations)) %>% 
  group_by(student_id) %>%
  pivot_wider(
    .,
    names_from = course_id,
    values_from = grade
  ) %>%
  ungroup() %>% 
  rename(Specialization = specialisation, ID = student_id)
data_wide
```

***

# Descriptives

```{r counting}
# Table of students per picked specialisation
full_tidy_data %>%
  filter(!is.na(specialisation)) %>% 
  distinct(student_id, .keep_all = T) %>% 
  mutate(
    specialisation = if_else(
      specialisation %in% c("0", "Algemene Psychologie"),
      "Did not (yet) pick a specialisation",
      specialisation
    )
  ) %>% 
  group_by(specialisation, cohort) %>% 
  count() %>% 
  pivot_wider(specialisation, names_from = cohort, values_from = n) %>% 
  mutate(
    Total = `1718` + `1819`
  ) %>%
  rename(
    Specialisation = specialisation,
    `17-18` = `1718`,
    `18-19` = `1819`
  ) %>% 
  knitr::kable()

full_tidy_data %>% 
  filter(cohort == "1920" | !specialisation %in% c("0", "Algemene Psychologie")) %>%
  distinct(student_id, .keep_all = T) %>% 
  ggplot(aes(x = cohort, fill = specialisation)) +
  geom_bar(position = "fill") +
  theme_minimal() +
  labs(
    x = "Cohort",
    y = "Fraction of Students",
    fill = "Specialisation"
  )
```

```{r grade_distribution}
ggplot(data, aes(x = grade)) +
  geom_bar(fill = "#f6a70a") +
  facet_wrap(vars(short_description)) +
  labs(
    x = "Grade",
    y = "No. of Students"
  )
```

***

# Network Analyses

## Exploratory Analyses using partial correlation networks

While exploring different options of adjusting grades for e.g. course or student means we discovered that course means differed only minimally, with a standard deviation of only $SD = `r sd(data %>% group_by(course_id, cohort) %>% summarise(grade = mean(grade, na.rm =T)) %>% pull(grade)) %>% round(3)`$ and that adjusting for course mean grades thus did not lead to any noticable differences in networks. Further, with only five first-year courses being present across the cohorts and their grades being highly correlated, adjusting for individual student's overall performance i.e. their GPA removed too much information from the different course's grades for networks to be estimated. Our initial estimated network is thus estimated using partial correlations between students' raw grades.

### Overall Network

```{r,fig.width = 10, fig.height = 10}
network <- estimateNetwork(data_wide[,-c(1:2)],
                           default = "pcor",
                           corMethod = "spearman")

qgraph(network$graph,
       layout = "circle",
       theme = "colorblind")
```

### Specialization Networks

#### Estimating a network for every specialization {.tabset}

##### Clinical
```{r,fig.width = 10, fig.height = 10}
# Estimate Network for All students who have specialized in Clinical
dataCLIN <- 
  data_wide %>%
      filter(data_wide$Specialization == "Spec Klinische Psychologie")


networkCLIN <- estimateNetwork(dataCLIN[,-c(1:2)],
                               default = "pcor",
                               corMethod = "spearman")

qgraph(networkCLIN$graph,
       layout = "circle",
       theme = "colorblind",
       title.cex = 2,
       title = paste("Spec Klinische Psychologie; N =", nrow(dataCLIN))) # N = 45
```

##### Work & Organisational
```{r,fig.width = 10, fig.height = 10}
# Estimate Network for All students who have specialized in Work & Organisational 
dataWOP <- 
  data_wide %>%
      filter(data_wide$Specialization == "Spec Arbeids- en Org. Psych.")


networkWOP <- estimateNetwork(dataWOP[,-c(1:2)],
                               default = "pcor",
                               corMethod = "spearman")

qgraph(networkWOP$graph,
       layout = "circle",
       theme = "colorblind",
       title.cex = 2,
       title = paste("Spec Arbeids- en Org. Psych.; N =", nrow(dataWOP))) # N = 64
```

##### Social
```{r, ,fig.width = 10, fig.height = 10}
# Estimate Network for All students who have specialized in Social
dataSOC <- 
  data_wide %>%
    filter(data_wide$Specialization == "Spec Sociale Psychologie")

networkSOC <- estimateNetwork(dataSOC[,-c(1:2)],
                               default = "pcor",
                               corMethod = "spearman")

qgraph(networkSOC$graph,
       layout = "circle",
       theme = "colorblind",
       title.cex = 2,
       title = paste("Spec Sociale Psychologie; N =", nrow(dataSOC))) # N = 56
```

##### Brain & Cognition
```{r, ,fig.width = 10, fig.height = 10}
# Estimate Network for All students who have specialized in Brain & Cognition
dataBC <- 
  data_wide %>%
      filter(data_wide$Specialization == "Spec Brein en Cognitie")


networkBC <- estimateNetwork(dataBC[,-c(1:2)],
                               default = "pcor",
                               corMethod = "spearman")

qgraph(networkBC$graph,
       layout = "circle",
       theme = "colorblind",
       title.cex = 2,
       title = paste("Spec Brein en Cognitie; N =", nrow(dataBC))) # N = 40
```

##### Clinical Dev
```{r,fig.width = 10, fig.height = 10}
# Estimate Network for All students who have specialized in Developmental
dataDEV <- 
  data_wide %>%
      filter(data_wide$Specialization == "Spec Klinische Ontwikkelingsps")

networkDEV <- estimateNetwork(dataDEV[,-c(1:2)],
                               default = "pcor",
                               corMethod = "spearman")

qgraph(networkDEV$graph,
       layout = "circle",
       theme = "colorblind",
       title.cex = 2,
       title = paste("Spec Klinische Ontwikkelingsps; N =", nrow(dataDEV))) # N = 50
```

##### Clinical Neuro
```{r,fig.width = 10, fig.height = 10}
# Estimate Network for All students who have specialized in Clinical Neuro
dataCNP <- 
  data_wide %>%
      filter(data_wide$Specialization == "Spec Klinische Neuropsych.")

networkCNP <- estimateNetwork(dataCNP[,-c(1:2)],
                               default = "pcor",
                               corMethod = "spearman")

qgraph(networkCNP$graph,
       layout = "circle",
       theme = "colorblind",
       title.cex = 2,
       title = paste("Spec Klinische Neuropsych.	; N =", nrow(dataCNP))) # N = 22
```

##### Methods
```{r, fig.width = 10, fig.height = 10}
# Estimate Network for All students who have specialized in Methods
dataPML <- 
  data_wide %>%
      filter(data_wide$Specialization == "Spec Psych. Methodenleer")

networkPML <- estimateNetwork(dataPML[,-c(1:2)],
                               default = "pcor",
                               corMethod = "spearman")

qgraph(networkPML$graph,
       layout = "circle",
       theme = "colorblind",
       title.cex = 2,
       title = paste("Spec Psych. Methodenleer; N =", nrow(dataPML))) # N = 31
```

#### Estimating a network for every specialization using a dummy (Yes/No) {.tabset}

##### Clinical
```{r,fig.width = 10, fig.height = 10}
# Create Dummy Variable for "CLIN" Specialization: Yes/No
dataCLIN <- 
  data_wide %>%
      mutate(CLIN = ifelse(data_wide$Specialization == "Spec Klinische Psychologie", 1, 0))


networkCLIN <- estimateNetwork(dataCLIN[,-c(1:2)],
                               default = "pcor",
                               corMethod = "spearman")

qgraph(networkCLIN$graph,
       layout = "circle",
       theme = "colorblind",
       title.cex = 2,
       title = paste("Spec Klinische Psychologie (Yes/No); N =", nrow(dataCLIN)))
```

##### Work & Organisational
```{r,fig.width = 10, fig.height = 10}
# Create Dummy Variable for "WOP" Specialization: Yes/No
dataWOP <- 
  data_wide %>%
      mutate(WOP = ifelse(data_wide$Specialization == "Spec Arbeids- en Org. Psych.", 1, 0))


networkWOP <- estimateNetwork(dataWOP[,-c(1:2)],
                               default = "pcor",
                               corMethod = "spearman")

qgraph(networkWOP$graph,
       layout = "circle",
       theme = "colorblind",
       title.cex = 2,
       title = paste("Spec Arbeids- en Org. Psych. (Yes/No); N =", nrow(dataWOP)))
```

##### Social
```{r,fig.width = 10, fig.height = 10}
# Create Dummy Variable for "SOC" Specialization: Yes/No
dataSOC <- 
  data_wide %>%
      mutate(SOC = ifelse(data_wide$Specialization == "Spec Sociale Psychologie", 1, 0))


networkSOC <- estimateNetwork(dataSOC[,-c(1:2)],
                               default = "pcor",
                               corMethod = "spearman")

qgraph(networkSOC$graph,
       layout = "circle",
       theme = "colorblind",
       title.cex = 2,
       title = paste("Spec Sociale Psychologie (Yes/No); N =", nrow(dataSOC)))
```

##### Brain & Cognition
```{r,fig.width = 10, fig.height = 10}
# Create Dummy Variable for "BC" Specialization: Yes/No
dataBC <- 
  data_wide %>%
      mutate(BC = ifelse(data_wide$Specialization == "Spec Brein en Cognitie", 1, 0))


networkBC <- estimateNetwork(dataBC[,-c(1:2)],
                               default = "pcor",
                               corMethod = "spearman")

qgraph(networkBC$graph,
       layout = "circle",
       theme = "colorblind",
       title.cex = 2,
       title = paste("Spec Brein en Cognitie (Yes/No); N =", nrow(dataBC)))
```

##### Clinical Dev
```{r,fig.width = 10, fig.height = 10}
# Create Dummy Variable for "DEV" Specialization: Yes/No
dataDEV <- 
  data_wide %>%
      mutate(CDV = ifelse(data_wide$Specialization == "Spec Klinische Ontwikkelingsps", 1, 0))


networkDEV <- estimateNetwork(dataDEV[,-c(1:2)],
                               default = "pcor",
                               corMethod = "spearman")

qgraph(networkDEV$graph,
       layout = "circle",
       theme = "colorblind",
       title.cex = 2,
       title = paste("Spec Klinische Ontwikkelingsps(Yes/No); N =", nrow(dataDEV)))
```

##### Clinical Neuro
```{r,fig.width = 10, fig.height = 10}
# Create Dummy Variable for "DEV" Specialization: Yes/No
dataCNP <- 
  data_wide %>%
      mutate(CNP = ifelse(data_wide$Specialization == "Spec Klinische Neuropsych.", 1, 0))


networkCNP <- estimateNetwork(dataCNP[,-c(1:2)],
                               default = "pcor",
                               corMethod = "spearman")

qgraph(networkCNP$graph,
       layout = "circle",
       theme = "colorblind",
       title.cex = 2,
       title = paste("Spec Klinische Neuropsych.(Yes/No); N =", nrow(dataCNP)))
```

##### Methods
```{r,fig.width = 10, fig.height = 10}
# Create Dummy Variable for "PML" Specialization: Yes/No
dataPML <- 
  data_wide %>%
      mutate(PML = ifelse(data_wide$Specialization == "Spec Psych. Methodenleer", 1, 0))


networkPML <- estimateNetwork(dataPML[,-c(1:2)],
                               default = "pcor",
                               corMethod = "spearman")

qgraph(networkPML$graph,
       layout = "circle",
       theme = "colorblind",
       title.cex = 2,
       title = paste("Spec Psych. Methodenleer (Yes/No); N =", nrow(dataPML)))
```

## Centrality and Stability

Centrality refers to the relative importance a variable has within the network. This relative importance derives from either the importance one variable has on the rest of the network, or from the way the variables within a network relate. 

### Overall Network

```{r, fig.width=10, fig.height=10}
network <- estimateNetwork(data_wide[,-c(1:2)],
                           default = "pcor",
                           corMethod = "spearman")
qgraph(network$graph,
       layout = "circle",
       theme = "colorblind")
```

```{r}
centralityPlot(network$graph,
               include = c("Strength",
                           "Closeness",
                           "Betweenness"))
```



```{r, cache = TRUE, message = FALSE, warning = FALSE}
nCores <- parallel::detectCores()

set.seed(222)

# Non-Parametric Bootstrap
bootNonParametric <- bootnet(network,
                             nBoots = 1000,
                             nCores = nCores - 1)
# Case-Drop Bootstrap
bootCaseDrop <- bootnet(network,
                        nBoots = 1000,
                        nCores = nCores - 1,
                        type = "case",
                        statistics = c("Strength", "Closeness", "Betweenness"))
```

```{r, cache = TRUE, message = FALSE, warning = FALSE}
# Plot Edge Accuracy Non-Parametric Bootstrap
plot(bootNonParametric,
     order = "sample",
     labels = FALSE)

# Plot Edge Differences Non-Parametric Bootstrap
plot(bootNonParametric,
     plot = "difference",
     onlyNonZero= TRUE,
     order = "sample")

# Plot Centrality Accuracy Case-Drop Bootstrap
plot(bootCaseDrop,
     statistics = c("Strength", "Closeness", "Betweenness"))

# CS-Coefficient
corStability(bootCaseDrop)

# Difference Test
plot(bootCaseDrop, "strength")
```

### Clinical vs Non-Clinical

```{r}
dataClin <- data_wide[data_wide$Specialization == "Spec Klinische Psychologie",]

dataNotClin <- data_wide[data_wide$Specialization != "Spec Klinische Psychologie",]
```

```{r}
networkCLIN <- estimateNetwork(dataClin[,-c(1:2)],
                               default = "pcor",
                               corMethod = "spearman")

networkNotCLIN <- estimateNetwork(dataNotClin[,-c(1:2)],
                               default = "pcor",
                               corMethod = "spearman")
```

```{r, fig.width=10, fig.height=10}
avgLayout <- averageLayout(networkCLIN, networkNotCLIN)

qgraph(networkCLIN$graph,
       layout = "circle",
       theme = "colorblind",
       title.cex = 2,
       title = paste("Spec Klinische Psychologie; N =", nrow(dataClin)))
qgraph(networkNotCLIN$graph,
       layout = "circle",
       theme = "colorblind",
       title.cex = 2,
       title = paste("Not Spec Klinische Psychologie; N =", nrow(dataNotClin)))
```

```{r}
centralityPlot(networkCLIN,
               include = c("Strength",
                           theme_bw = TRUE,
                           "Closeness",
                           "Betweenness"))

centralityPlot(networkNotCLIN,
               theme_bw = TRUE,
               include = c("Strength",
                           "Closeness",
                           "Betweenness"))

centralityPlot(list(Clinical = networkCLIN,
                    `Not Clinical` = networkNotCLIN),
               theme_bw = TRUE,
               include = c("Strength",
                           "Closeness",
                           "Betweenness"))
```

```{r, message = FALSE, warning = FALSE}
set.seed(45)
NCT <- 
NetworkComparisonTest::NCT(networkCLIN, networkNotCLIN,
                           it = 2000,
                           test.edges = TRUE,
                           edges = "all",
                           progressbar = FALSE)

NCT
NCT$einv.pvals %>% 
    filter(`p-value` <= .05)
```

```{r,cache = TRUE, message = FALSE, warning = FALSE}
nCores <- parallel::detectCores()

set.seed(555)

# Non-Parametric Bootstrap
bootNonParametricCLIN <- bootnet(networkCLIN,
                             nBoots = 1000,
                             nCores = nCores - 1)
# Case-Drop Bootstrap
bootCaseDropCLIN <- bootnet(networkCLIN,
                        nBoots = 1000,
                        nCores = nCores - 1,
                        type = "case",
                        statistics = c("Strength", "Closeness", "Betweenness"))
```

```{r, message = FALSE, warning = FALSE}
# Plot Edge Accuracy Non-Parametric Bootstrap
plot(bootNonParametricCLIN,
     order = "sample",
     labels = FALSE)

# Plot Edge Differences Non-Parametric Bootstrap
plot(bootNonParametricCLIN,
     plot = "difference",
     onlyNonZero= TRUE,
     order = "sample")

# Plot Centrality Accuracy Case-Drop Bootstrap
plot(bootCaseDropCLIN,
     statistics = c("Strength", "Closeness", "Betweenness"))

# CS-Coefficient
corStability(bootCaseDropCLIN)

# Difference Test
plot(bootNonParametricCLIN, "strength")
```

```{r, cache = TRUE, message = FALSE, warning = FALSE}
nCores <- parallel::detectCores()

set.seed(444)

# Non-Parametric Bootstrap
bootNonParametricNotCLIN <- bootnet(networkNotCLIN,
                             nBoots = 1000,
                             nCores = nCores - 1)
# Case-Drop Bootstrap
bootCaseDropNotCLIN<- bootnet(networkNotCLIN,
                        nBoots = 1000,
                        nCores = nCores - 1,
                        type = "case",
                        statistics = c("Strength", "Closeness", "Betweenness"))
```

```{r, message = FALSE, warning = FALSE}
# Plot Edge Accuracy Non-Parametric Bootstrap
plot(bootNonParametricNotCLIN,
     order = "sample",
     labels = FALSE)

# Plot Edge Differences Non-Parametric Bootstrap
plot(bootNonParametricNotCLIN,
     plot = "difference",
     onlyNonZero= TRUE,
     order = "sample")

# Plot Centrality Accuracy Case-Drop Bootstrap
plot(bootCaseDropNotCLIN,
     statistics = c("Strength", "Closeness", "Betweenness"))

# CS-Coefficient
corStability(bootCaseDropNotCLIN)

# Difference Test
plot(bootNonParametricNotCLIN, "strength")
```

## Mixed Graphical Models (MGMs)

```{r mgm_setup}
library(mgm)

# Transform long grade data into wide format
generate_wide_grades <- function (data, grade_column) {
  data %>%
    mutate(
      course_abb = recode(course_id, !!!course_abbrevations)
    ) %>% 
    pivot_wider(
      id_cols = student_id,
      names_from = course_abb,
      
      # Use variable to assign grade column
      values_from = grade_column
    ) %>% 
    # Add a column with specialisation info
    left_join(
      data %>%
        select(student_id, specialisation) %>%
        distinct(),
      by = "student_id"
    )
}

# Collect some info about columns
# This just needs to be a list of characters for each column indicating
# the type. Automated code here is just for flexibility.
extract_mgm_column_info <- . %>% 
  # Get only TRUE/FALSE whether columns are numeric
  summarise(across(everything(), is.numeric)) %>% 
  # Get into long format with column names and whether they are numeric in 2 columns
  pivot_longer(
    everything(),
    names_to = "column",
    values_to = "is_numeric"
  ) %>% 
  # Assign letters based on whether column is numeric or not
  mutate(
    column = recode(column, specialisation = "SPEC"),
    # Types for mgm
    # g: gaussian (normal dist.)
    # c: categorical
    # p: poisson (if skewed)
    type = ifelse(is_numeric, "g", "c")
  )
```

### MGM Predicting all Specializations at Once

We first fit a mixed graphical model with contiuous gaussian nodes for the course grades and a single categorical node for the different specializations.

```{r mgm_numerical, fig.height=12, fig.width=12}
set.seed(2021)
# ==== Preparate data for MGM ====
mgm_num_data <- data %>% 
  generate_wide_grades("grade") %>%
  # Drop Student IDs
  select(-student_id) %>% 
  # Drop all NAs (NOTE: this loses us about 2/3 of the data)
  na.omit()

column_info <- extract_mgm_column_info(mgm_num_data)

# Transform categorical variables to integers
# mgm wants all data as numbers!
mgm_num_data_numeric <- mgm_num_data %>% 
  mutate(
    specialisation = as.numeric(as.factor(specialisation))
  )

# V1: mgm package directly
trainidx <- sample(seq_len(nrow(mgm_num_data)), size = floor(.8 * nrow(mgm_num_data)))

fit_mgm_num <- mgm(
  data = mgm_num_data_numeric[trainidx,],
  type = column_info$type
)

# Predict the training data
prediction_mgm <- predict(
  fit_mgm_num,
  data = mgm_num_data_numeric[-trainidx,],
  errorCat = c("CC","nCC","CCmarg"),
  errorCon = c("R2")
)

 # Prediction error
prediction_mgm$errors %>% knitr::kable()

# Unfortunately this model merely predicts every student to
# pick the clinical specialization
prediction_mgm$predicted %>% head()

# Obtain data to highlight prediction accuracy in network graph
error_list <- list()
color_list <- list()
for (i in 1:nrow(column_info)) {
  if (column_info$type[[i]] == "g") {
    error_list[[i]] <- prediction_mgm$errors[i, 2]
    color_list[[i]] <- "#90B4D4" 
    
  } else if (column_info$type[[i]] == "c") {
    beyondmarg <- prediction_mgm$errors[i, 3] - prediction_mgm$errors[i, 5]
    error_list[[i]] <- c(prediction_mgm$errors[i, 5],beyondmarg)
    
    color_list[[i]] <- c("#ffa500", "#ff4300")
    
  } else {
    stop("Unsupported column type")
  }
}

# Plot the network graph
qgraph(
  fit_mgm_num$pairwise$wadj,
  pie = error_list,
  layout="circle",
  labels = column_info$column,
  pieColor = color_list,
  label.cex = .9,
  curveAll = TRUE,
  curveDefault = .6,
  cut = 0,
  theme = "colorblind"
)
```

### MGM Predicting Clinical vs Non-Clinical

```{r}
# Create data frame excluding student IDs
dataMGM <- data_wide[,-2]

# Recode specialization to numeric 
dataMGM %<>% 
  mutate(Clinical = ifelse(Specialization %in% "Spec Klinische Psychologie", 1, 0)) %>%
    select(-Specialization)

# Create sample for prediction
set.seed(25)
trainidx <- sample((seq_len(nrow(dataMGM))), size = floor(.8 * nrow(dataMGM)))
```

```{r}
# Fitting a MGM with the mgm-package
set.seed(1)
fit_obj <- mgm(data = na.omit(dataMGM[trainidx,]), # na.omit excludes N = 32 
                type = c(rep("g", 5),"c"),
                level = c(rep(1, 5), 2),
                ruleReg = "OR",
                k = 2,
                binarySign = TRUE)
```

```{r}
# Node Predictability estimation
p_obj <- predict(fit_obj, na.omit(dataMGM[-trainidx,]),
                  errorCat = c("CC","; NCC","CCmarg"),
                  errorCon = c("R2"))

p_obj$errors

p_obj$predicted[,6] %>%
  table()

dataMGM[-trainidx,] %>%
  na.omit() %>%
    select(Clinical) %>%
      table() %>%
        proportions()
```

```{r}
error_list <- list() # List for ring-segments
for(i in 1:5)
  error_list[[i]] <- p_obj$errors[i,2]
  beyondmarg <- p_obj$errors[6,3]-p_obj$errors[6,5]
  error_list[[6]] <- c(p_obj$errors[6,5],beyondmarg)

color_list <- list() # List for Colors

for(i in 1:5)
  color_list[[i]] <- "#90B4D4"
  color_list[[6]] <- c("#ffa500", "#ff4300")
```

```{r, fig.width=12, fig.height=12}
set.seed(10)
qgraph(fit_obj$pairwise$wadj,
       pie = error_list,
       layout="circle",
       theme = "colorblind",
       labels = colnames(dataMGM),
       pieColor = color_list,
       label.cex = .9,
       curveAll = TRUE,
       curveDefault = .6,
       cut = 0,
       labels = colnames(dataMGM))
```

*** 

# Multinomial Prediction

## Examining regression features

```{r}
# Construct wide data frame to fit a multinomial model
data_multinom <- data_wide %>% 
  mutate(
    Specialization = relevel(as.factor(Specialization), ref = "Spec Klinische Psychologie")
  ) %>% 
  left_join(
    data %>%
      select(student_id, cohort),
    by = c(ID = "student_id")
  ) %>% 
  select(-ID)

# Order alphabetically
spec_order <- data_multinom$Specialization %>%
  unique() %>%
  as.character()

# add features to data 
data_multinom_feature <- data_multinom
feature_matrix <- matrix(NA, nrow = nrow(data_multinom), ncol = 3)
feature_matrix[,1] <- apply(data_multinom[,2:6],1, max)
feature_matrix[,2] <- apply(data_multinom[,2:6],1, min)
feature_matrix[,3] <- apply(data_multinom[,2:6],1, mean)
data_multinom_feature <- cbind(data_multinom_feature, feature_matrix)

# Fit the multinomial feature data only on data from the 1718 cohort 
fitted_1718_features <- multinom(Specialization ~ ., data = data_multinom_feature %>% filter(cohort == "1718") %>% select(-cohort))


# Try to predict the 1819 cohort we're using the raw probabilities here
# because assigning to the highest probability will
# always pick the clinical specialization
prediction <- predict(fitted_1718_features, newdata = data_multinom_feature %>% filter(cohort == "1819") %>% select(-cohort), type = "probs")

# Get predicted fraction of people
frac_prediction <- prediction %>%
  colMeans(na.rm = T) 

# Get observed fraction of people for 1718
frac_1718 <- data %>%
  distinct(student_id, .keep_all = T) %>%
  select(specialisation, cohort) %>% 
  filter(cohort == "1718") %>% 
  count(specialisation) %>% 
  mutate(freq = n / sum(n)) %>%
  pull(freq, name = specialisation)

# Get observed fraction of people for 1819
frac_1819 <- data %>%
  distinct(student_id, .keep_all = T) %>%
  select(specialisation, cohort) %>% 
  filter(cohort == "1819") %>% 
  count(specialisation) %>% 
  mutate(freq = n / sum(n)) %>%
  pull(freq, name = specialisation)

# Examine prediction accuracy via correlations
paste("Correlation using last year's proportions:", cor(frac_1718[spec_order], frac_1819[spec_order]))
paste("Correlation using multinomial predictions:", cor(frac_prediction[spec_order], frac_1819[spec_order]))

# Examine raw prediction accuracy and visualise it in a plot
data.frame(
  specialisation = spec_order,
  Correct = frac_1819[spec_order],
  Predicted = frac_prediction[spec_order],
  Last_Year = frac_1718[spec_order]
) %>%
  pivot_longer(-specialisation) %>% 
  ggplot(aes(x = factor(name, levels = c("Last_Year", "Predicted", "Correct")), y = value, fill = specialisation)) +
    geom_bar(position = "fill", stat = "identity") +
    theme_minimal() +
    labs(
      x = NULL,
      y = "Fraction of Students",
      fill = "Specialisation"
    )
```




## Evaluation of multinomial performance
```{r multinom_eval}
# Fit the multinomial only on data from the 1718 cohort
fitted_1718 <- multinom(Specialization ~ ., data = data_multinom %>% filter(cohort == "1718") %>% select(-cohort))

# Try to predict the 1819 cohort we're using the raw probabilities here
# because assigning to the highest probability will
# always pick the clinical specialization
prediction <- predict(fitted_1718, newdata = data_multinom %>% filter(cohort == "1819") %>% select(-cohort), type = "probs")

# Get predicted fraction of people
frac_prediction <- prediction %>%
  colMeans(na.rm = T) 

# Get observed fraction of people for 1718
frac_1718 <- data %>%
  distinct(student_id, .keep_all = T) %>%
  select(specialisation, cohort) %>% 
  filter(cohort == "1718") %>% 
  count(specialisation) %>% 
  mutate(freq = n / sum(n)) %>%
  pull(freq, name = specialisation)

# Get observed fraction of people for 1819
frac_1819 <- data %>%
  distinct(student_id, .keep_all = T) %>%
  select(specialisation, cohort) %>% 
  filter(cohort == "1819") %>% 
  count(specialisation) %>% 
  mutate(freq = n / sum(n)) %>%
  pull(freq, name = specialisation)

# Examine prediction accuracy via correlations
paste("Correlation using last year's proportions:", cor(frac_1718[spec_order], frac_1819[spec_order]))
paste("Correlation using multinomial predictions:", cor(frac_prediction[spec_order], frac_1819[spec_order]))

# Examine raw prediction accuracy and visualise it in a plot
data.frame(
  specialisation = spec_order,
  Correct = frac_1819[spec_order],
  Predicted = frac_prediction[spec_order],
  Last_Year = frac_1718[spec_order]
) %>%
  pivot_longer(-specialisation) %>% 
  ggplot(aes(x = factor(name, levels = c("Last_Year", "Predicted", "Correct")), y = value, fill = specialisation)) +
    geom_bar(position = "fill", stat = "identity") +
    theme_minimal() +
    labs(
      x = NULL,
      y = "Fraction of Students",
      fill = "Specialisation"
    )
```

