---
title: "Report (Final version, Jan's partial)"
author: "Wessley Kameraad, René Freichel, Jan Simson"
date: "18.01.2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
	echo = FALSE,
	message = FALSE,
	warning = FALSE
)
```

```{r, echo = FALSE, message = FALSE, warning = FALSE}
installed_packages <- installed.packages()[, "Package"]

## Install Missing Required Packages
if (!"tidyverse" %in% installed_packages) {
  install.packages("tidyverse")
  }
if (!"qgraph" %in% installed_packages) {
  install.packages("qgraph")
}
if (!"dplyr" %in% installed_packages) {
  install.packages("dplyr")
}
if (!"magrittr" %in% installed_packages) {
  install.packages("magrittr")
}
if (!"bootnet" %in% installed_packages) {
  install.packages("bootnet")
}
if (!"mgm" %in% installed_packages) {
  install.packages("mgm")
}
if (!"janitor" %in% installed_packages) {
  install.packages("janitor")
}
if (!requireNamespace("BiocManager", quietly = TRUE)){
  install.packages("BiocManager")
  BiocManager::install("Rgraphviz")
}
    

## Load packages
library("tidyverse")
library("qgraph")
library("dplyr")
library("magrittr")
library("bootnet")
library("mgm")
library("janitor")
library("bnlearn")
library("Rgraphviz")
library("nnet")

set.seed(2021)
```

```{r load_raw_data, message=FALSE, warning=FALSE}
cohorts <- c("1718", "1819", "1920")

# Load and combine the three CSVs
raw_data <- data.frame()
for (cohort in cohorts) {
  file <- paste0("data/anonymized_cohort", cohort, ".csv")
  
  cohort_data <- read_csv2(file) %>%
    clean_names() %>% 
    mutate(
      # Add a new column "cohort"
      cohort = cohort
    )

  
  # Add new cohorts data to joined data frame
  raw_data <- bind_rows(
    raw_data,
    cohort_data
  )
}
```

```{r tidy_data}
# Create a tidy dataframe with a subset of the data
full_tidy_data <- raw_data %>% 
  transmute(
    # Parse the grade as a decimal number
    grade = parse_number(
      offic_cijfer,
      locale = locale(decimal_mark = ",", grouping_mark = ".")
    ) %>%
      suppressWarnings,
    # Put all special grades (i.e. letters) into a separate column
    grade_special = if_else(is.na(grade), offic_cijfer, as.character(NA)),
    # Rename & select other columns
    specialisation = specialisatie,
    course_id = studiegids,
    student_id = masked_id,
    # The short description has different column names in different cohorts
    short_description = if_else(is.na(omschr_10), omschr_11, omschr_10),
    long_description = curs_omschr_l,
    cohort = cohort,
    rank = rangnummer,
    country = land
  )
```

```{r filter_data}
selected_courses <- c(
  "7201602PXY", # 1. Introduction Psychology & Cognition
  "7201604PXY", # 2. Research Methods & Statistics
  "7201605PXY", # 3. Developmental Psy
  "7201607PXY", # 4. Social & Work Psy
  "7201622PXY" # 5. Clinical Psychology & Brain
  # 5. is matched with "7201610PXY": Clinical & Biological Psy (1718)
)

# Merge the two courses across cohorts:
# Clinical Psychology & Brain (cohorts 1819, 1920)
# Clinical & Biological Psy (cohort 1718)
# and remove any instances where a student has taken both courses
# (that means they failed the first one, so that is dropped)
students_with_double_bio_courses <- full_tidy_data %>%
  filter(course_id %in% c("7201622PXY", "7201610PXY")) %>% 
  count(student_id) %>% 
  filter(n > 1) %>% 
  pull(student_id)
full_tidy_data <- full_tidy_data %>%
  filter(
    # Only keep the newer bio course if a student has both
    # (this means they failed to complete the old bio course)
    course_id != "7201610PXY" | !(student_id %in% students_with_double_bio_courses)
  ) %>% 
  mutate(
    course_id = if_else(course_id == "7201610PXY", "7201622PXY", course_id),
    short_description = if_else(
      course_id == "7201622PXY",
      "Clinical Psychology & Brain / Bio",
      short_description
    ),
    long_description = if_else(
      course_id == "7201622PXY", 
      "[Combined] Clinical Psychology & Brain + Clinical & Biological Psy", 
      long_description
    )
  )

data <- full_tidy_data %>%
  # Only include students who picked a specialisation
  filter(!is.na(specialisation) & specialisation != "0") %>% 
  filter(specialisation != "Algemene Psychologie") %>% 
  # Include only first year courses
  filter(course_id %in% selected_courses)
```

```{r}
course_abbrevations <- c(
  "7201602PXY" = "IPC",
  "7201604PXY" = "RMS",
  "7201605PXY" = "DEV",
  "7201607PXY" = "SWO",
  "7201622PXY" = "CB"
)

data_wide <- data %>%
  select(specialisation, grade, course_id, student_id) %>%
  mutate(course_id = recode(course_id, !!!course_abbrevations)) %>% 
  group_by(student_id) %>%
  pivot_wider(
    .,
    names_from = course_id,
    values_from = grade
  ) %>%
  ungroup() %>% 
  rename(Specialization = specialisation, ID = student_id)
data_wide
```

## Descriptives

```{r counting}
# Table of students per picked specialisation
full_tidy_data %>%
  filter(!is.na(specialisation)) %>% 
  distinct(student_id, .keep_all = T) %>% 
  mutate(
    specialisation = if_else(
      specialisation %in% c("0", "Algemene Psychologie"),
      "Did not (yet) pick a specialisation",
      specialisation
    )
  ) %>% 
  group_by(specialisation, cohort) %>% 
  count() %>% 
  pivot_wider(specialisation, names_from = cohort, values_from = n) %>% 
  mutate(
    Total = `1718` + `1819`
  ) %>%
  rename(
    Specialisation = specialisation,
    `17-18` = `1718`,
    `18-19` = `1819`
  ) %>% 
  knitr::kable()

full_tidy_data %>% 
  filter(cohort == "1920" | !specialisation %in% c("0", "Algemene Psychologie")) %>%
  distinct(student_id, .keep_all = T) %>% 
  ggplot(aes(x = cohort, fill = specialisation)) +
  geom_bar(position = "fill") +
  theme_minimal() +
  labs(
    x = "Cohort",
    y = "Fraction of Students",
    fill = "Specialisation"
  )
```

```{r grade_distribution}
ggplot(data, aes(x = grade)) +
  geom_bar(fill = "#f6a70a") +
  facet_wrap(vars(short_description)) +
  labs(
    x = "Grade",
    y = "No. of Students"
  )
```


***

## Gaussian Networks

While exploring different options of adjusting grades for e.g. course or student means we discovered that course means differed only minimally, with a standard deviation of only $SD = `r sd(data %>% group_by(course_id, cohort) %>% summarise(grade = mean(grade, na.rm =T)) %>% pull(grade)) %>% round(3)`$ and that adjusting for course mean grades thus did not lead to any noticable differences in networks. Further, with only five first-year courses being present across the cohorts and their grades being highly correlated, adjusting for individual student's overall performance i.e. their GPA removed too much information from the different course's grades for networks to be estimated. Our initial estimated network is thus estimated using partial correlations between students' raw grades.

```{r}
networkGrades <- data_wide %>%
  select(-Specialization, -ID) %>% 
  estimateNetwork(
    default = "pcor",
    threshold = "sig",
    alpha = .05
  )

qgraph(networkGrades$graph)
```

***

## Mixed Graphical Models (MGM)

```{r mgm_setup}
library(mgm)

# Transform long grade data into wide format
generate_wide_grades <- function (data, grade_column) {
  data %>%
    mutate(
      course_abb = recode(course_id, !!!course_abbrevations)
    ) %>% 
    pivot_wider(
      id_cols = student_id,
      names_from = course_abb,
      
      # Use variable to assign grade column
      values_from = grade_column
    ) %>% 
    # Add a column with specialisation info
    left_join(
      data %>%
        select(student_id, specialisation) %>%
        distinct(),
      by = "student_id"
    )
}

# Collect some info about columns
# This just needs to be a list of characters for each column indicating
# the type. Automated code here is just for flexibility.
extract_mgm_column_info <- . %>% 
  # Get only TRUE/FALSE whether columns are numeric
  summarise(across(everything(), is.numeric)) %>% 
  # Get into long format with column names and whether they are numeric in 2 columns
  pivot_longer(
    everything(),
    names_to = "column",
    values_to = "is_numeric"
  ) %>% 
  # Assign letters based on whether column is numeric or not
  mutate(
    column = recode(column, specialisation = "SPEC"),
    # Types for mgm
    # g: gaussian (normal dist.)
    # c: categorical
    # p: poisson (if skewed)
    type = ifelse(is_numeric, "g", "c")
  )
```

### Mixed Graphical Model using all specializations

We first fit a mixed graphical model with contiuous gaussian nodes for the course grades and a single categorical node for the different specializations.

```{r mgm_numerical, fig.height=12, fig.width=12}
set.seed(2021)
# ==== Preparate data for MGM ====
mgm_num_data <- data %>% 
  generate_wide_grades("grade") %>%
  # Drop Student IDs
  select(-student_id) %>% 
  # Drop all NAs (NOTE: this loses us about 2/3 of the data)
  na.omit()

column_info <- extract_mgm_column_info(mgm_num_data)

# Transform categorical variables to integers
# mgm wants all data as numbers!
mgm_num_data_numeric <- mgm_num_data %>% 
  mutate(
    specialisation = as.numeric(as.factor(specialisation))
  )

# V1: mgm package directly
trainidx <- sample(seq_len(nrow(mgm_num_data)), size = floor(.8 * nrow(mgm_num_data)))

fit_mgm_num <- mgm(
  data = mgm_num_data_numeric[trainidx,],
  type = column_info$type
)

# Predict the training data
prediction_mgm <- predict(
  fit_mgm_num,
  data = mgm_num_data_numeric[-trainidx,],
  errorCat = c("CC","nCC","CCmarg"),
  errorCon = c("R2")
)

 # Prediction error
prediction_mgm$errors %>% knitr::kable()

# Unfortunately this model merely predicts every student to
# pick the clinical specialization
prediction_mgm$predicted %>% head()

# Obtain data to highlight prediction accuracy in network graph
error_list <- list()
color_list <- list()
for (i in 1:nrow(column_info)) {
  if (column_info$type[[i]] == "g") {
    error_list[[i]] <- prediction_mgm$errors[i, 2]
    color_list[[i]] <- "#90B4D4" 
    
  } else if (column_info$type[[i]] == "c") {
    beyondmarg <- prediction_mgm$errors[i, 3] - prediction_mgm$errors[i, 5]
    error_list[[i]] <- c(prediction_mgm$errors[i, 5],beyondmarg)
    
    color_list[[i]] <- c("#ffa500", "#ff4300")
    
  } else {
    stop("Unsupported column type")
  }
}

# Plot the network graph
qgraph(
  fit_mgm_num$pairwise$wadj,
  pie = error_list,
  layout="circle",
  labels = column_info$column,
  pieColor = color_list,
  label.cex = .9,
  curveAll = TRUE,
  curveDefault = .6,
  cut = 0,
  theme = "colorblind"
)
```

### Experimental: Predicting the 1920 cohort using the MGM.

We also explored predicting specializations for the 1920 cohort using the MGM fitted above. However, the model predicted all students to pick the clinical specialization as well and thus we abandoned this approach.

```{r mgm_num_predict_1920}
new_mgm_num_data <- full_tidy_data %>% 
  filter(cohort == 1920) %>% 
  # Include only first year courses
  filter(course_id %in% selected_courses) %>% 
  generate_wide_grades("grade") %>% 
  select(-student_id) %>% 
  # Just put in random numbers for specialisation
  mutate(specialisation = sample(1:7, nrow(.), replace = T)) %>% 
  na.omit()

new_prediction_mgm <- predict(fit_mgm_num, data = new_mgm_num_data)
```

*** 

## Multinomial Prediction

> TODO: Insert multinomial code by René here

### Evaluation of multinomial performance:
```{r multinom_eval}
# Construct wide data frame to fit a multinomial model
data_multinom <- data_wide %>% 
  mutate(
    Specialization = relevel(as.factor(Specialization), ref = "Spec Klinische Psychologie")
  ) %>% 
  left_join(
    data %>%
      select(student_id, cohort),
    by = c(ID = "student_id")
  ) %>% 
  select(-ID)

# Order alphabetically
spec_order <- data_multinom$Specialization %>%
  unique() %>%
  as.character()

# Fit the multinomial only on data from the 1718 cohort
fitted_1718 <- multinom(Specialization ~ ., data = data_multinom %>% filter(cohort == "1718") %>% select(-cohort))

# Try to predict the 1819 cohort we're using the raw probabilities here
# because assigning to the highest probability will
# always pick the clinical specialization
prediction <- predict(fitted_1718, newdata = data_multinom %>% filter(cohort == "1819") %>% select(-cohort), type = "probs")

# Get predicted fraction of people
frac_prediction <- prediction %>%
  colMeans(na.rm = T) 

# Get observed fraction of people for 1718
frac_1718 <- data %>%
  distinct(student_id, .keep_all = T) %>%
  select(specialisation, cohort) %>% 
  filter(cohort == "1718") %>% 
  count(specialisation) %>% 
  mutate(freq = n / sum(n)) %>%
  pull(freq, name = specialisation)

# Get observed fraction of people for 1819
frac_1819 <- data %>%
  distinct(student_id, .keep_all = T) %>%
  select(specialisation, cohort) %>% 
  filter(cohort == "1819") %>% 
  count(specialisation) %>% 
  mutate(freq = n / sum(n)) %>%
  pull(freq, name = specialisation)

# Examine prediction accuracy via correlations
paste("Correlation using last year's proportions:", cor(frac_1718[spec_order], frac_1819[spec_order]))
paste("Correlation using multinomial predictions:", cor(frac_prediction[spec_order], frac_1819[spec_order]))

# Examine raw prediction accuracy and visualise it in a plot
data.frame(
  specialisation = spec_order,
  Correct = frac_1819[spec_order],
  Predicted = frac_prediction[spec_order],
  Last_Year = frac_1718[spec_order]
) %>%
  pivot_longer(-specialisation) %>% 
  ggplot(aes(x = factor(name, levels = c("Last_Year", "Predicted", "Correct")), y = value, fill = specialisation)) +
    geom_bar(position = "fill", stat = "identity") +
    theme_minimal() +
    labs(
      x = NULL,
      y = "Fraction of Students",
      fill = "Specialisation"
    )
```

