---
title: "N-check"
author: "Wesley Kammeraad"
date: "21-1-2021"
output: html_document
---
 
# Jan 
 
```{r load_raw_data}
library(tidyverse)
library(janitor)

cohorts <- c("1718", "1819", "1920")

# Load and combine the three CSVs
raw_data <- data.frame()
for (cohort in cohorts) {
  file <- paste0("data/anonymized_cohort", cohort, ".csv")
  
  cohort_data <- read_csv2(file) %>%
    clean_names() %>% 
    mutate(
      # Add a new column "cohort"
      cohort = cohort
    )

  
  # Add new cohorts data to joined data frame
  raw_data <- bind_rows(
    raw_data,
    cohort_data
  )
}
```

```{r tidy_data}
# Create a tidy dataframe with a subset of the data
full_tidy_data <- raw_data %>% 
  transmute(
    # Parse the grade as a decimal number
    grade = parse_number(
      offic_cijfer,
      locale = locale(decimal_mark = ",", grouping_mark = ".")
    ) %>%
      suppressWarnings,
    # Put all special grades (i.e. letters) into a separate column
    grade_special = if_else(is.na(grade), offic_cijfer, as.character(NA)),
    # Rename & select other columns
    specialisation = specialisatie,
    course_id = studiegids,
    student_id = masked_id,
    # The short description has different column names in different cohorts
    short_description = if_else(is.na(omschr_10), omschr_11, omschr_10),
    long_description = curs_omschr_l,
    cohort = cohort,
    rank = rangnummer,
    country = land
  )
```

```{r filter_data}
selected_courses <- c(
  "7201602PXY", # 1. Introduction Psychology & Cognition
  "7201604PXY", # 2. Research Methods & Statistics
  "7201605PXY", # 3. Developmental Psy
  "7201607PXY", # 4. Social & Work Psy
  "7201610PXY" # 5. Clinical & Biological Psy
)

data <- full_tidy_data %>%
  # Only include students who picked a specialisation
  filter(!is.na(specialisation) & specialisation != "0") %>% 
  filter(specialisation != "Algemene Psychologie") %>% 
  # Include only first year courses
  filter(course_id %in% selected_courses)
```

```{r}
data %<>%
  select(specialisation, student_id, grade, course_id) %>%
  pivot_wider(.,
              names_from = course_id,
              values_from = grade)

data
```

```{r}
# Assigning uniform column names for the courses
colnames(data) <-
  c("Specialization",
    "ID",
    "IP&C",
    "RM&S",
    "DevPsy",
    "S&WO",
    "C&B")

data
```

```{r}
nJan <- 
data %>%
  distinct(ID, Specialization) %>%
    count(Specialization)

nrowJan <- nrow(data)
```

----------------------------------------------------------------------------------------------------------------------------------------------------------------------

# Wesley

```{r, echo = FALSE, message = FALSE, warning = FALSE}
installed_packages <- installed.packages()[, "Package"]

## Install Missing Required Packages
if (!"tidyverse" %in% installed_packages) {
  install.packages("tidyverse")
  }
if (!"qgraph" %in% installed_packages) {
  install.packages("qgraph")
}
if (!"dplyr" %in% installed_packages) {
  install.packages("dplyr")
}
if (!"psych" %in% installed_packages) {
  install.packages("magrittr")
}
if (!"bootnet" %in% installed_packages) {
  install.packages("bootnet")
}
if (!"mgm" %in% installed_packages) {
  install.packages("mgm")
}

## Load packages
library("tidyverse")
library("qgraph")
library("dplyr")
library("magrittr")
library("bootnet")
library("mgm")
```

## %in%; Study Guide Number

```{r}
data1718 <- read.csv("data/anonymized_cohort1718.csv",
                     header = TRUE,
                     sep = ";",
                     dec = ",",
                     na.strings = c("AVV", "-"))

data1819 <- read.csv("data/anonymized_cohort1819.csv",
                     header = TRUE,
                     sep = ";",
                     dec = ",",
                     na.strings = c("AVV", "-"))

data1920 <- read.csv("data/anonymized_cohort1920.csv",
                     header = TRUE,
                     sep = ";",
                     dec = ",",
                     na.strings = c("AVV", "-"))
```

```{r}
colnames(data1718) <- c("Track", "Subject", "Country", "Specialization", "Grade", "Course", "CourseDesc", "Study Guide", "ID")

colnames(data1819) <- c("Track", "Subject", "Country", "Specialization", "Grade", "Course", "CourseDesc", "Study Guide", "Rank", "ID")

colnames(data1920) <- c("Track", "Subject", "Country", "Grade", "Course", "CourseDesc", "Study Guide", "Rank", "ID")  
```

```{r}
# Changing grades from character to numeric data
data1718$Grade <- gsub(",", ".", data1718$Grade)
data1718$Grade <- as.numeric(data1718$Grade)

data1819$Grade <- gsub(",", ".", data1819$Grade)
data1819$Grade <- as.numeric(data1819$Grade)
```

```{r}
data1718 %>%
  select(CourseDesc, `Study Guide`) %>%
    unique()

data1718 %>%
  select(`Study Guide`) %>%
    unique()


data1718 %>%
  select(CourseDesc) %>%
    unique()
```
Shared Program 1718

> Introductory Psychology and Cognition	7201602PXY			
> Research Methods and Statistics	7201604PXY			
> Developmental Psychology	7201605PXY			
> Social and Work & Organisational Psychology	7201607PXY			
> Clinical Psychology & Brain and Cognition	7201622PXY

Shared Program 1819

> Introductory Psychology and Cognition	7201602PXY			 
> Research Methods and Statistics	7201604PXY			
> Developmental Psychology	7201605PXY			
> Social and Work & Organisational Psychology	7201607PXY			
> Clinical Psychology & Brain and Cognition	7201622PXY

```{r}
# Retain Shared Program Courses
courses1718 <- 
  data1718 %>%
    select(CourseDesc, `Study Guide`) %>%
      unique() %>%
        filter(CourseDesc %in% c("Introductory Psychology and Cognition",
                               "Research Methods and Statistics",
                               "Developmental Psychology",
                               "Social and Work & Organisational Psychology",
                               "Clinical and Biological Psychology"))

courses1819 <- 
  data1819 %>%
    select(CourseDesc, `Study Guide`) %>%
      unique() %>%
        filter(CourseDesc %in% c("Introductory Psychology and Cognition",
                               "Research Methods and Statistics",
                               "Developmental Psychology",
                               "Social and Work & Organisational Psychology",
                               "Clinical Psychology & Brain and Cognition"))
```

```{r}
data1718 <- data1718[data1718$`Study Guide` %in% courses1718$`Study Guide`,]
data1819 <- data1819[data1819$`Study Guide` %in% courses1819$`Study Guide`,]
```

```{r}
data1718 %<>%
  select(Specialization, Grade, Course, ID) %>%
  pivot_wider(.,
              names_from = Course,
              values_from = Grade)

data1819 %<>%
  select(Specialization, Grade, Course, ID) %>%
  pivot_wider(.,
              names_from = Course,
              values_from = Grade)
```

```{r}
# Assigning uniform column names for the courses
colnames(data1718) <-
  colnames(data1819) <-
  c("Specialization",
    "ID",
    "IP&C",
    "RM&S",
    "DevPsy",
    "S&WO",
    "C&B")

data1718
data1819
```

```{r}
data <- rbind(data1718, data1819)
data
```

```{r}
data %>%
  distinct(ID, Specialization) %>%
    count(Specialization)

nrow(data)
```

```{r}
# Removing all students that have not picked a specialization
dataSpec <- data[data$Specialization != 0,]
dataSpec <- dataSpec[dataSpec$Specialization != "Algemene Psychologie",]
dataSpec
```

```{r}
nWes <- 
dataSpec %>%
  distinct(ID, Specialization) %>%
    count(Specialization)

nrowWes <- nrow(dataSpec)
```


```{r, confirmation}
nJan == nWes
nrowJan == nrowWes
``` 
