---
title: "Report"
author: "Wessley Kameraad, René Freichel, Jan Simson"
date: "18.01.2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, echo = FALSE, message = FALSE, warning = FALSE}
installed_packages <- installed.packages()[, "Package"]

## Install Missing Required Packages
if (!"tidyverse" %in% installed_packages) {
  install.packages("tidyverse")
  }
if (!"qgraph" %in% installed_packages) {
  install.packages("qgraph")
}
if (!"dplyr" %in% installed_packages) {
  install.packages("dplyr")
}
if (!"magrittr" %in% installed_packages) {
  install.packages("magrittr")
}
if (!"bootnet" %in% installed_packages) {
  install.packages("bootnet")
}
if (!"mgm" %in% installed_packages) {
  install.packages("mgm")
}
if (!"janitor" %in% installed_packages) {
  install.packages("janitor")
}
if (!requireNamespace("BiocManager", quietly = TRUE)){
  install.packages("BiocManager")
  BiocManager::install("Rgraphviz")
}
    

## Load packages
library("tidyverse")
library("qgraph")
library("dplyr")
library("magrittr")
library("bootnet")
library("mgm")
library("janitor")
library("bnlearn")
library("Rgraphviz")

set.seed(2021)
```

```{r load_raw_data, message=FALSE, warning=FALSE}
cohorts <- c("1718", "1819", "1920")

# Load and combine the three CSVs
raw_data <- data.frame()
for (cohort in cohorts) {
  file <- paste0("data/anonymized_cohort", cohort, ".csv")
  
  cohort_data <- read_csv2(file) %>%
    clean_names() %>% 
    mutate(
      # Add a new column "cohort"
      cohort = cohort
    )

  
  # Add new cohorts data to joined data frame
  raw_data <- bind_rows(
    raw_data,
    cohort_data
  )
}
```

```{r tidy_data}
# Create a tidy dataframe with a subset of the data
full_tidy_data <- raw_data %>% 
  transmute(
    # Parse the grade as a decimal number
    grade = parse_number(
      offic_cijfer,
      locale = locale(decimal_mark = ",", grouping_mark = ".")
    ) %>%
      suppressWarnings,
    # Put all special grades (i.e. letters) into a separate column
    grade_special = if_else(is.na(grade), offic_cijfer, as.character(NA)),
    # Rename & select other columns
    specialisation = specialisatie,
    course_id = studiegids,
    student_id = masked_id,
    # The short description has different column names in different cohorts
    short_description = if_else(is.na(omschr_10), omschr_11, omschr_10),
    long_description = curs_omschr_l,
    cohort = cohort,
    rank = rangnummer,
    country = land
  )
```

```{r filter_data}
selected_courses <- c(
  "7201602PXY", # 1. Introduction Psychology & Cognition
  "7201604PXY", # 2. Research Methods & Statistics
  "7201605PXY", # 3. Developmental Psy
  "7201607PXY", # 4. Social & Work Psy
  "7201622PXY" # 5. Clinical Psychology & Brain
  # 5. is matched with "7201610PXY": Clinical & Biological Psy (1718)
)

# Merge the two courses across cohorts:
# Clinical Psychology & Brain (cohorts 1819, 1920)
# Clinical & Biological Psy (cohort 1718)
# and remove any instances where a student has taken both courses
# (that means they failed the first one, so that is dropped)
students_with_double_bio_courses <- full_tidy_data %>%
  filter(course_id %in% c("7201622PXY", "7201610PXY")) %>% 
  count(student_id) %>% 
  filter(n > 1) %>% 
  pull(student_id)
full_tidy_data <- full_tidy_data %>%
  filter(
    # Only keep the newer bio course if a student has both
    # (this means they failed to complete the old bio course)
    course_id != "7201610PXY" | !(student_id %in% students_with_double_bio_courses)
  ) %>% 
  mutate(
    course_id = if_else(course_id == "7201610PXY", "7201622PXY", course_id),
    short_description = if_else(
      course_id == "7201622PXY",
      "Clinical Psychology & Brain / Bio",
      short_description
    ),
    long_description = if_else(
      course_id == "7201622PXY", 
      "[Combined] Clinical Psychology & Brain + Clinical & Biological Psy", 
      long_description
    )
  )

data <- full_tidy_data %>%
  # Only include students who picked a specialisation
  filter(!is.na(specialisation) & specialisation != "0") %>% 
  filter(specialisation != "Algemene Psychologie") %>% 
  # Include only first year courses
  filter(course_id %in% selected_courses)
```

```{r}
course_abbrevations <- c(
  "7201602PXY" = "IPC",
  "7201604PXY" = "RMS",
  "7201605PXY" = "DEV",
  "7201607PXY" = "SWO",
  "7201622PXY" = "CB"
)

data_wide <- data %>%
  select(specialisation, grade, course_id, student_id) %>%
  mutate(course_id = recode(course_id, !!!course_abbrevations)) %>% 
  group_by(student_id) %>%
  pivot_wider(
    .,
    names_from = course_id,
    values_from = grade
  ) %>%
  ungroup() %>% 
  rename(Specialization = specialisation, ID = student_id)
data_wide
```

```{r counting}
# Table of students per picked specialisation
full_tidy_data %>%
  filter(!is.na(specialisation)) %>% 
  distinct(student_id, .keep_all = T) %>% 
  mutate(
    specialisation = if_else(
      specialisation %in% c("0", "Algemene Psychologie"),
      "Did not (yet) pick a specialisation",
      specialisation
    )
  ) %>% 
  group_by(specialisation, cohort) %>% 
  count() %>% 
  pivot_wider(specialisation, names_from = cohort, values_from = n) %>% 
  mutate(
    Total = `1718` + `1819`
  ) %>%
  rename(
    Specialisation = specialisation,
    `17-18` = `1718`,
    `18-19` = `1819`
  ) %>% 
  knitr::kable()

full_tidy_data %>% 
  filter(cohort == "1920" | !specialisation %in% c("0", "Algemene Psychologie")) %>%
  distinct(student_id, .keep_all = T) %>% 
  ggplot(aes(x = cohort, fill = specialisation)) +
  geom_bar(position = "fill") +
  theme_minimal() +
  labs(
    x = "Cohort",
    y = "Fraction of Students",
    fill = "Specialisation"
  )
```
***

## Gaussian Networks

While exploring different options of adjusting grades for e.g. course or student means we discovered that course means differed only minimally, with a standard deviation of only $SD = `r sd(data %>% group_by(course_id, cohort) %>% summarise(grade = mean(grade, na.rm =T)) %>% pull(grade)) %>% round(3)`$ and that adjusting for course mean grades thus did not lead to differences in networks. Further, with only five first-year courses being consistent across cohorts, adjusting for individual student's overall performance removed too much information from the individual course's grades that networks could not be estimated anymore. Our initial estimated network is thus estimated using student's raw grades.

```{r}
networkGrades <- data_wide %>%
  select(-Specialization, -ID) %>% 
  estimateNetwork(
    default = "pcor",
    threshold = "sig",
    alpha = .05
  )

qgraph(networkGrades$graph)
```

***

## Mixed Graphical Models

```{r mgm_setup}
library(mgm)

generate_wide_grades <- function (data, grade_column) {
  data %>%
    mutate(
      course_abb = recode(course_id, !!!course_abbrevations)
    ) %>% 
    pivot_wider(
      id_cols = student_id,
      names_from = course_abb,
      
      # Use variable to assign grade column
      values_from = grade_column
    ) %>% 
    # Add a column with specialisation info
    left_join(
      data %>%
        select(student_id, specialisation) %>%
        distinct(),
      by = "student_id"
    )
}

# Collect some info about columns
# This just needs to be a list of characters for each column indicating
# the type. Automated code here is just for flexibility.
extract_mgm_column_info <- . %>% 
  # Get only TRUE/FALSE whether columns are numeric
  summarise(across(everything(), is.numeric)) %>% 
  # Get into long format with column names and whether they are numeric in 2 columns
  pivot_longer(
    everything(),
    names_to = "column",
    values_to = "is_numeric"
  ) %>% 
  # Assign letters based on whether column is numeric or not
  mutate(
    column = recode(column, specialisation = "SPEC"),
    # Types for mgm
    # g: gaussian (normal dist.)
    # c: categorical
    # p: poisson (if skewed)
    type = ifelse(is_numeric, "g", "c")
  )
```

### Initial Mixed Graphical Model

Fit a mgm using gaussian distributions for grade data, this has approach is more interpretable, but has a much lower N due to NAs in the data which have to be dropped per row.
```{r mgm_numerical}
# ==== Preparate data for MGM ====
mgm_num_data <- data %>% 
  generate_wide_grades("grade") %>%
  # Drop Student IDs
  select(-student_id) %>% 
  # Drop all NAs (NOTE: this loses us about 2/3 of the data)
  na.omit()

column_info <- extract_mgm_column_info(mgm_num_data)

# Transform categorical variables to integers
# mgm wants all data as numbers!
mgm_num_data_numeric <- mgm_num_data %>% 
  mutate(
    specialisation = as.numeric(as.factor(specialisation))
  )

# V1: mgm package directly
fit_mgm_num <- mgm(
  data = mgm_num_data_numeric,
  type = column_info$type
)
qgraph::qgraph(
  fit_mgm_num$pairwise$wadj,
  labels = column_info$column,
  theme = "colorblind",
  layout = "spring"
)

# Create a factor graph (whatever that is?)
# Also shows interaction effects as squares
FactorGraph(fit_mgm_num)

# Predict the training data
prediction_mgm <- predict(
  fit_mgm_num,
  data = mgm_num_data_numeric,
  errorCat = c("CC","nCC","CCmarg"),
  errorCon = c("R2")
)
prediction_mgm$errors # Prediction error
prediction_mgm$predicted %>% head()

# Construct a fancy pie graph
error_list <- list()
color_list <- list()
for (i in 1:nrow(column_info)) {
  if (column_info$type[[i]] == "g") {
    error_list[[i]] <- prediction_mgm$errors[i, 2]
    color_list[[i]] <- "#90B4D4" 
    
  } else if (column_info$type[[i]] == "c") {
    beyondmarg <- prediction_mgm$errors[i, 3] - prediction_mgm$errors[i, 5]
    error_list[[i]] <- c(prediction_mgm$errors[i, 5],beyondmarg)
    
    color_list[[i]] <- c("#ffa500", "#ff4300")
    
  } else {
    stop("Unsupported column type")
  }
}

qgraph(
  fit_mgm_num$pairwise$wadj,
   pie = error_list,
   layout="spring",
   labels = column_info$column,
   pieColor = color_list,
   label.cex = .9,
   edge.color = fit_mgm_num$pairwise$edgecolor,
   curveAll = TRUE,
   curveDefault = .6,
   cut = 0
)
```

### Predict the 1920 cohort using the numerical mgm.

> Note: we might want to drop this part because currently predictions are all category 5: clinical.

```{r mgm_num_predict_1920}
new_mgm_num_data <- full_tidy_data %>% 
  filter(cohort == 1920) %>% 
  # Include only first year courses
  filter(course_id %in% selected_courses) %>% 
  generate_wide_grades("grade") %>% 
  select(-student_id) %>% 
  # Just put in random numbers for specialisation
  mutate(specialisation = sample(1:7, nrow(.), replace = T)) %>% 
  na.omit()

new_prediction_mgm <- predict(fit_mgm_num, data = new_mgm_num_data)
new_prediction_mgm$predicted %>% head(20)
```

***

### Trichotomised Split 

Here we bin the data according to grade:

* < 6 = 1; These are the NAs, which imply failed grades
* 6 - 7 = 2
* > 7 = 3

```{r}
dataOrdinal <- data_wide
dataOrdinal[is.na(dataOrdinal)] <- 0

dataOrdinal %<>%
    select("IPC", "RMS", "DEV","SWO", "CB") %>% 
    lapply(., function(x){ifelse(x < 6, 1,
                                 ifelse(x > 7, 3, 2))}) %>%
        as.data.frame()

networkOrdinal <- estimateNetwork(dataOrdinal,
                                   default = "pcor",
                                   corMethod = "cor_auto")

qgraph(networkOrdinal$graph)
```

---------------------------------------------------------------------------------------------------------------

### Bayesian Network

```{r}
dataBN <- na.omit(data_wide[,-c(1, 2)])
estBN <- hc(dataBN,
            restart = 10,
            perturb = 100)

estBN
```

```{r}
(edgeStrength <- arc.strength(estBN, dataBN, "bic-g"))
```

```{r}
strength.plot(estBN,
              edgeStrength,
              main = "Bayesian Network",
              shape = "ellipse")
```

```{r}
set.seed(333)
bootBN <- boot.strength(dataBN,
                        R = 2000,
                        algorithm = "hc")
```

```{r}
averageBN <- averaged.network(bootBN,
                              threshold = .95) # Edges appear in 95% of estimated networks

edgeStrengthAvg <- arc.strength(averageBN,
                                dataBN,
                                "bic-g")

strength.plot(averageBN,
              edgeStrengthAvg,
              main = "Bootstrapped Bayesian",
              shape = "ellipse")
```

```{r}
subEdge <- bootBN[bootBN$strength > 0.95,]
subEdge
```

```{r}
bootTab <- bootBN[bootBN$strength > 0.85 &
                    bootBN$direction > 0.5,]
avgStr <- bootTab
avgStr$strength <- bootTab$direction

strength.plot(averageBN,
              avgStr,
              main = "Averaged Bayesian Network",
              shape = "ellipse")
```

---------------------------------------------------------------------------------------------------------------
### Full Information Maximum Likelihood Estimation

```{r}
# Create column for number of NAs in row
dataImpute <- data_wide
dataImpute$NAs <- apply(dataImpute, 1, function(x){sum(is.na(x))})
```

```{r}
# Full Information Max Likelihood for Cases with one (1) missing value
networkImputeNA1 <- 
  # Retain rows with one or less missing value
  dataImpute[dataImpute$NAs < 2,] %>% 
    # Only retain the course grade columns
    select(c("IPC", "RMS", "DEV","SWO", "CB")) %>%  
      estimateNetwork(.,
                      default = "EBICglasso",
                      corMethod = "cor_auto",
                      missing = "fiml")
# Estimate graph
qgraph(networkImputeNA1$graph,
       layout = "spring",
       theme = "colorblind")
```
