---
title: "Wesley_Final-Report-Code"
author: "Wesley Kammeraad"
date: "27-1-2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, echo = FALSE, message = FALSE, warning = FALSE}
installed_packages <- installed.packages()[, "Package"]

## Install Missing Required Packages
if (!"tidyverse" %in% installed_packages) {
  install.packages("tidyverse")
  }
if (!"qgraph" %in% installed_packages) {
  install.packages("qgraph")
}
if (!"dplyr" %in% installed_packages) {
  install.packages("dplyr")
}
if (!"magrittr" %in% installed_packages) {
  install.packages("magrittr")
}
if (!"bootnet" %in% installed_packages) {
  install.packages("bootnet")
}
if (!"mgm" %in% installed_packages) {
  install.packages("mgm")
}
if (!"janitor" %in% installed_packages) {
  install.packages("janitor")
}
if (!"Rgraphviz" %in% installed_packages) {
  if (!"BiocManager" %in% installed_packages){
    install.packages("BiocManager")
  }
  BiocManager::install("Rgraphviz")
}
    
## Load All Required Packages
library("tidyverse")
library("qgraph")
library("dplyr")
library("magrittr")
library("bootnet")
library("mgm")
library("janitor")
library("bnlearn")
library("Rgraphviz")
```

```{r load_raw_data, message=FALSE, warning=FALSE}
cohorts <- c("1718", "1819", "1920")

# Load and combine the three CSVs
raw_data <- data.frame()
for (cohort in cohorts) {
  file <- paste0("data/anonymized_cohort", cohort, ".csv")
  
  cohort_data <- read_csv2(file) %>%
    clean_names() %>% 
    mutate(
      # Add a new column "cohort"
      cohort = cohort
    )

  
  # Add new cohorts data to joined data frame
  raw_data <- bind_rows(
    raw_data,
    cohort_data
  )
}
```


```{r tidy_data}
# Create a tidy dataframe with a subset of the data
full_tidy_data <- raw_data %>% 
  transmute(
    # Parse the grade as a decimal number
    grade = parse_number(
      offic_cijfer,
      locale = locale(decimal_mark = ",", grouping_mark = ".")
    ) %>%
      suppressWarnings,
    # Put all special grades (i.e. letters) into a separate column
    grade_special = if_else(is.na(grade), offic_cijfer, as.character(NA)),
    # Rename & select other columns
    specialisation = specialisatie,
    course_id = studiegids,
    student_id = masked_id,
    # The short description has different column names in different cohorts
    short_description = if_else(is.na(omschr_10), omschr_11, omschr_10),
    long_description = curs_omschr_l,
    cohort = cohort,
    rank = rangnummer,
    country = land
  )
```

```{r filter_data}
selected_courses <- c(
  "7201602PXY", # 1. Introduction Psychology & Cognition
  "7201604PXY", # 2. Research Methods & Statistics
  "7201605PXY", # 3. Developmental Psy
  "7201607PXY", # 4. Social & Work Psy
  "7201622PXY" # 5. Clinical Psychology & Brain
  # 5. is matched with "7201610PXY": Clinical & Biological Psy (1718)
)

# Merge the two courses across cohorts:
# Clinical Psychology & Brain (cohorts 1819, 1920)
# Clinical & Biological Psy (cohort 1718)
# and remove any instances where a student has taken both courses
# (that means they failed the first one, so that is dropped)
students_with_double_bio_courses <- full_tidy_data %>%
  filter(course_id %in% c("7201622PXY", "7201610PXY")) %>% 
  count(student_id) %>% 
  filter(n > 1) %>% 
  pull(student_id)
full_tidy_data <- full_tidy_data %>%
  filter(
    # Only keep the newer bio course if a student has both
    # (this means they failed to complete the old bio course)
    course_id != "7201610PXY" | !(student_id %in% students_with_double_bio_courses)
  ) %>% 
  mutate(
    course_id = if_else(course_id == "7201610PXY", "7201622PXY", course_id),
    short_description = if_else(
      course_id == "7201622PXY",
      "Clinical Psychology & Brain / Bio",
      short_description
    ),
    long_description = if_else(
      course_id == "7201622PXY", 
      "[Combined] Clinical Psychology & Brain + Clinical & Biological Psy", 
      long_description
    )
  )

data <- full_tidy_data %>%
  # Only include students who picked a specialisation
  filter(!is.na(specialisation) & specialisation != "0") %>% 
  filter(specialisation != "Algemene Psychologie") %>% 
  # Include only first year courses
  filter(course_id %in% selected_courses)
```

```{r}
data_wide <- data %>%
  select(specialisation, grade, course_id, student_id) %>%
    group_by(student_id) %>%
      pivot_wider(.,
                  names_from = course_id,
                  values_from = grade) %>%
          ungroup()
data_wide
```

```{r}
# Assigning uniform column names for the courses
colnames(data_wide) <-
  c("Specialization",
    "ID",
    "IPC",              #7201602PXY
    "RMS",              #7201604PXY
    "DEV",              #7201605PXY
    "SWO",              #7201607PXY
    "CB")               #7201610PXY
```

```{r}
data_wide %>%
  count(Specialization) %>%
   mutate(total = sum(n),
          proportion = round(proportions(n), 2))

```

***

### Network Analysis

#### Overall Network

```{r,fig.width = 10, fig.height = 10}
network <- estimateNetwork(data_wide[,-c(1:2)],
                           default = "pcor",
                           corMethod = "spearman")

qgraph(network$graph,
       layout = "circle",
       theme = "colorblind")
```

#### Estimating a network for every specialization

```{r,fig.width = 10, fig.height = 10}
# Estimate Network for All students who have specialized in Clinical
dataCLIN <- 
  data_wide %>%
      filter(data_wide$Specialization == "Spec Klinische Psychologie")


networkCLIN <- estimateNetwork(dataCLIN[,-c(1:2)],
                               default = "pcor",
                               corMethod = "spearman")

qgraph(networkCLIN$graph,
       layout = "circle",
       theme = "colorblind",
       title.cex = 2,
       title = paste("Spec Klinische Psychologie; N =", nrow(dataCLIN))) # N = 45
```

```{r,fig.width = 10, fig.height = 10}
# Estimate Network for All students who have specialized in Work & Organisational 
dataWOP <- 
  data_wide %>%
      filter(data_wide$Specialization == "Spec Arbeids- en Org. Psych.")


networkWOP <- estimateNetwork(dataWOP[,-c(1:2)],
                               default = "pcor",
                               corMethod = "spearman")

qgraph(networkWOP$graph,
       layout = "circle",
       theme = "colorblind",
       title.cex = 2,
       title = paste("Spec Arbeids- en Org. Psych.; N =", nrow(dataWOP))) # N = 64
```

```{r, ,fig.width = 10, fig.height = 10}
# Estimate Network for All students who have specialized in Social
dataSOC <- 
  data_wide %>%
    filter(data_wide$Specialization == "Spec Sociale Psychologie")

networkSOC <- estimateNetwork(dataSOC[,-c(1:2)],
                               default = "pcor",
                               corMethod = "spearman")

qgraph(networkSOC$graph,
       layout = "circle",
       theme = "colorblind",
       title.cex = 2,
       title = paste("Spec Sociale Psychologie; N =", nrow(dataSOC))) # N = 56
```

```{r, ,fig.width = 10, fig.height = 10}
# Estimate Network for All students who have specialized in Brain & Cognition
dataBC <- 
  data_wide %>%
      filter(data_wide$Specialization == "Spec Brein en Cognitie")


networkBC <- estimateNetwork(dataBC[,-c(1:2)],
                               default = "pcor",
                               corMethod = "spearman")

qgraph(networkBC$graph,
       layout = "circle",
       theme = "colorblind",
       title.cex = 2,
       title = paste("Spec Brein en Cognitie; N =", nrow(dataBC))) # N = 40
```

```{r,fig.width = 10, fig.height = 10}
# Estimate Network for All students who have specialized in Developmental
dataDEV <- 
  data_wide %>%
      filter(data_wide$Specialization == "Spec Klinische Ontwikkelingsps")

networkDEV <- estimateNetwork(dataDEV[,-c(1:2)],
                               default = "pcor",
                               corMethod = "spearman")

qgraph(networkDEV$graph,
       layout = "circle",
       theme = "colorblind",
       title.cex = 2,
       title = paste("Spec Klinische Ontwikkelingsps; N =", nrow(dataDEV))) # N = 50
```

```{r,fig.width = 10, fig.height = 10}
# Estimate Network for All students who have specialized in Developmental
dataCNP <- 
  data_wide %>%
      filter(data_wide$Specialization == "Spec Klinische Neuropsych.")

networkCNP <- estimateNetwork(dataCNP[,-c(1:2)],
                               default = "pcor",
                               corMethod = "spearman")

qgraph(networkCNP$graph,
       layout = "circle",
       theme = "colorblind",
       title.cex = 2,
       title = paste("Spec Klinische Neuropsych.	; N =", nrow(dataCNP))) # N = 22
```

```{r, fig.width = 10, fig.height = 10}
# Estimate Network for All students who have specialized in Methods
dataPML <- 
  data_wide %>%
      filter(data_wide$Specialization == "Spec Psych. Methodenleer")

networkPML <- estimateNetwork(dataPML[,-c(1:2)],
                               default = "pcor",
                               corMethod = "spearman")

qgraph(networkPML$graph,
       layout = "circle",
       theme = "colorblind",
       title.cex = 2,
       title = paste("Spec Psych. Methodenleer; N =", nrow(dataPML))) # N = 31
```

##### Estimating a network for every specialization using a dummy (Yes/No)

```{r,fig.width = 10, fig.height = 10}
# Create Dummy Variable for "CLIN" Specialization: Yes/No
dataCLIN <- 
  data_wide %>%
      mutate(CLIN = ifelse(data_wide$Specialization == "Spec Klinische Psychologie", 1, 0))


networkCLIN <- estimateNetwork(dataCLIN[,-c(1:2)],
                               default = "pcor",
                               corMethod = "spearman")

qgraph(networkCLIN$graph,
       layout = "circle",
       theme = "colorblind",
       title.cex = 2,
       title = paste("Spec Klinische Psychologie (Yes/No); N =", nrow(dataCLIN)))
```

```{r,fig.width = 10, fig.height = 10}
# Create Dummy Variable for "WOP" Specialization: Yes/No
dataWOP <- 
  data_wide %>%
      mutate(WOP = ifelse(data_wide$Specialization == "Spec Arbeids- en Org. Psych.", 1, 0))


networkWOP <- estimateNetwork(dataWOP[,-c(1:2)],
                               default = "pcor",
                               corMethod = "spearman")

qgraph(networkWOP$graph,
       layout = "circle",
       theme = "colorblind",
       title.cex = 2,
       title = paste("Spec Arbeids- en Org. Psych. (Yes/No); N =", nrow(dataWOP)))
```

```{r,fig.width = 10, fig.height = 10}
# Create Dummy Variable for "SOC" Specialization: Yes/No
dataSOC <- 
  data_wide %>%
      mutate(SOC = ifelse(data_wide$Specialization == "Spec Sociale Psychologie", 1, 0))


networkSOC <- estimateNetwork(dataSOC[,-c(1:2)],
                               default = "pcor",
                               corMethod = "spearman")

qgraph(networkSOC$graph,
       layout = "circle",
       theme = "colorblind",
       title.cex = 2,
       title = paste("Spec Sociale Psychologie (Yes/No); N =", nrow(dataSOC)))
```

```{r,fig.width = 10, fig.height = 10}
# Create Dummy Variable for "BC" Specialization: Yes/No
dataBC <- 
  data_wide %>%
      mutate(BC = ifelse(data_wide$Specialization == "Spec Brein en Cognitie", 1, 0))


networkBC <- estimateNetwork(dataBC[,-c(1:2)],
                               default = "pcor",
                               corMethod = "spearman")

qgraph(networkBC$graph,
       layout = "circle",
       theme = "colorblind",
       title.cex = 2,
       title = paste("Spec Brein en Cognitie (Yes/No); N =", nrow(dataBC)))
```

```{r,fig.width = 10, fig.height = 10}
# Create Dummy Variable for "DEV" Specialization: Yes/No
dataDEV <- 
  data_wide %>%
      mutate(CDV = ifelse(data_wide$Specialization == "Spec Klinische Ontwikkelingsps", 1, 0))


networkDEV <- estimateNetwork(dataDEV[,-c(1:2)],
                               default = "pcor",
                               corMethod = "spearman")

qgraph(networkDEV$graph,
       layout = "circle",
       theme = "colorblind",
       title.cex = 2,
       title = paste("Spec Klinische Ontwikkelingsps(Yes/No); N =", nrow(dataDEV)))
```

```{r,fig.width = 10, fig.height = 10}
# Create Dummy Variable for "DEV" Specialization: Yes/No
dataCNP <- 
  data_wide %>%
      mutate(CNP = ifelse(data_wide$Specialization == "Spec Klinische Neuropsych.", 1, 0))


networkCNP <- estimateNetwork(dataCNP[,-c(1:2)],
                               default = "pcor",
                               corMethod = "spearman")

qgraph(networkCNP$graph,
       layout = "circle",
       theme = "colorblind",
       title.cex = 2,
       title = paste("Spec Klinische Neuropsych.(Yes/No); N =", nrow(dataCNP)))
```

```{r,fig.width = 10, fig.height = 10}
# Create Dummy Variable for "PML" Specialization: Yes/No
dataPML <- 
  data_wide %>%
      mutate(PML = ifelse(data_wide$Specialization == "Spec Psych. Methodenleer", 1, 0))


networkPML <- estimateNetwork(dataPML[,-c(1:2)],
                               default = "pcor",
                               corMethod = "spearman")

qgraph(networkPML$graph,
       layout = "circle",
       theme = "colorblind",
       title.cex = 2,
       title = paste("Spec Psych. Methodenleer (Yes/No); N =", nrow(dataPML)))
```

### Centrality and Stability

#### Stability Overall Network: Centrality & Structure

```{r, fig.width=10, fig.height=10}
network <- estimateNetwork(data_wide[,-c(1:2)],
                           default = "pcor",
                           corMethod = "spearman")
qgraph(network$graph,
       layout = "circle",
       theme = "colorblind")
```

```{r}
centralityPlot(network$graph,
               include = c("Strength",
                           "Closeness",
                           "Betweenness"))
```



```{r, cache = TRUE, message = FALSE, warning = FALSE}
nCores <- parallel::detectCores()

set.seed(222)

# Non-Parametric Bootstrap
bootNonParametric <- bootnet(network,
                             nBoots = 1000,
                             nCores = nCores - 1)
# Case-Drop Bootstrap
bootCaseDrop <- bootnet(network,
                        nBoots = 1000,
                        nCores = nCores - 1,
                        type = "case",
                        statistics = c("Strength", "Closeness", "Betweenness"))
```

```{r, message = FALSE, warning = FALSE}
# Plot Edge Accuracy Non-Parametric Bootstrap
plot(bootNonParametric,
     order = "sample",
     labels = FALSE)

# Plot Edge Differences Non-Parametric Bootstrap
plot(bootNonParametric,
     plot = "difference",
     onlyNonZero= TRUE,
     order = "sample")

# Plot Centrality Accuracy Case-Drop Bootstrap
plot(bootCaseDrop,
     statistics = c("Strength", "Closeness", "Betweenness"))

# CS-Coefficient
corStability(bootCaseDrop)

# Difference Test
plot(bootCaseDrop, "strength")
```

#### Clinical vs Other Specs

```{r}
dataClin <- data_wide[data_wide$Specialization == "Spec Klinische Psychologie",]

dataNotClin <- data_wide[data_wide$Specialization != "Spec Klinische Psychologie",]
```

```{r}
networkCLIN <- estimateNetwork(dataClin[,-c(1:2)],
                               default = "pcor",
                               corMethod = "spearman")

networkNotCLIN <- estimateNetwork(dataNotClin[,-c(1:2)],
                               default = "pcor",
                               corMethod = "spearman")
```

```{r, fig.width=10, fig.height=10}
avgLayout <- averageLayout(networkCLIN, networkNotCLIN)

qgraph(networkCLIN$graph,
       layout = "circle",
       theme = "colorblind",
       title.cex = 2,
       title = paste("Spec Klinische Psychologie; N =", nrow(dataClin)))
qgraph(networkNotCLIN$graph,
       layout = "circle",
       theme = "colorblind",
       title.cex = 2,
       title = paste("Not Spec Klinische Psychologie; N =", nrow(dataNotClin)))
```

```{r}
centralityPlot(networkCLIN,
               include = c("Strength",
                           theme_bw = TRUE,
                           "Closeness",
                           "Betweenness"))

centralityPlot(networkNotCLIN,
               theme_bw = TRUE,
               include = c("Strength",
                           "Closeness",
                           "Betweenness"))

centralityPlot(list(Clinical = networkCLIN,
                    `Not Clinical` = networkNotCLIN),
               theme_bw = TRUE,
               include = c("Strength",
                           "Closeness",
                           "Betweenness"))
```

```{r, message = FALSE, warning = FALSE}
set.seed(45)
NCT <- 
NetworkComparisonTest::NCT(networkCLIN, networkNotCLIN,
                           it = 2000,
                           test.edges = TRUE,
                           edges = "all")

NCT
NCT$einv.pvals %>% 
    filter(`p-value` <= .05)
```

```{r,cache = TRUE, message = FALSE, warning = FALSE}
nCores <- parallel::detectCores()

set.seed(555)

# Non-Parametric Bootstrap
bootNonParametricCLIN <- bootnet(networkCLIN,
                             nBoots = 1000,
                             nCores = nCores - 1)
# Case-Drop Bootstrap
bootCaseDropCLIN <- bootnet(networkCLIN,
                        nBoots = 1000,
                        nCores = nCores - 1,
                        type = "case",
                        statistics = c("Strength", "Closeness", "Betweenness"))
```

```{r, message = FALSE, warning = FALSE}
# Plot Edge Accuracy Non-Parametric Bootstrap
plot(bootNonParametricCLIN,
     order = "sample",
     labels = FALSE)

# Plot Edge Differences Non-Parametric Bootstrap
plot(bootNonParametricCLIN,
     plot = "difference",
     onlyNonZero= TRUE,
     order = "sample")

# Plot Centrality Accuracy Case-Drop Bootstrap
plot(bootCaseDropCLIN,
     statistics = c("Strength", "Closeness", "Betweenness"))

# CS-Coefficient
corStability(bootCaseDropCLIN)

# Difference Test
plot(bootNonParametricCLIN, "strength")
```

```{r, cache = TRUE, message = FALSE, warning = FALSE}
nCores <- parallel::detectCores()

set.seed(444)

# Non-Parametric Bootstrap
bootNonParametricNotCLIN <- bootnet(networkNotCLIN,
                             nBoots = 1000,
                             nCores = nCores - 1)
# Case-Drop Bootstrap
bootCaseDropNotCLIN<- bootnet(networkNotCLIN,
                        nBoots = 1000,
                        nCores = nCores - 1,
                        type = "case",
                        statistics = c("Strength", "Closeness", "Betweenness"))
```

```{r, message = FALSE, warning = FALSE}
# Plot Edge Accuracy Non-Parametric Bootstrap
plot(bootNonParametricNotCLIN,
     order = "sample",
     labels = FALSE)

# Plot Edge Differences Non-Parametric Bootstrap
plot(bootNonParametricNotCLIN,
     plot = "difference",
     onlyNonZero= TRUE,
     order = "sample")

# Plot Centrality Accuracy Case-Drop Bootstrap
plot(bootCaseDropNotCLIN,
     statistics = c("Strength", "Closeness", "Betweenness"))

# CS-Coefficient
corStability(bootCaseDropNotCLIN)

# Difference Test
plot(bootNonParametricNotCLIN, "strength")
```

#### Clinical vs Non-Clinical Specs

```{r}
dataClin2 <- 
  data_wide[data_wide$Specialization %in%
              c("Spec Klinische Psychologie",
                "Spec Klinische Ontwikkelingsps",
                "Spec Klinische Neuropsych.",
                "Spec Brein en Cognitie"),]

dataNotClin2 <-
  data_wide[!data_wide$Specialization %in%
              c("Spec Klinische Psychologie",
                "Spec Klinische Ontwikkelingsps",
                "Spec Klinische Neuropsych.",
                "Spec Brein en Cognitie"),]
```

```{r}
networkCLIN2 <- estimateNetwork(dataClin2[,-c(1:2)],
                               default = "pcor",
                               corMethod = "spearman")

networkNotCLIN2 <- estimateNetwork(dataNotClin2[,-c(1:2)],
                               default = "pcor",
                               corMethod = "spearman")
```

```{r}
avgLayout2 <- averageLayout(networkCLIN2, networkNotCLIN2)

qgraph(networkCLIN2$graph,
       layout = avgLayout2,
       theme = "colorblind",
       title = paste("; N =", nrow(dataClin2)))
qgraph(networkNotCLIN2$graph,
       layout = avgLayout2,
       theme = "colorblind",
       title = paste("; N =", nrow(dataNotClin2)))
```

```{r, message = FALSE, warning = FALSE}
NCT2 <- 
NetworkComparisonTest::NCT(networkCLIN2, networkNotCLIN2,
                           it = 2000,
                           test.edges = TRUE,
                           edges = "all")

NCT2
NCT2$einv.pvals %>% 
    filter(`p-value` <= .05)
```

```{r}
centralityPlot(networkCLIN2,
               include = c("Strength",
                           theme_bw = TRUE,
                           "Closeness",
                           "Betweenness"))

centralityPlot(networkNotCLIN2,
               theme_bw = TRUE,
               include = c("Strength",
                           "Closeness",
                           "Betweenness"))


centralityPlot(list(Clinical = networkCLIN2,
                    `Not Clinical` = networkNotCLIN2),
               include = c("Strength",
                           theme_bw = TRUE,
                           "Closeness",
                           "Betweenness"))
```

```{r, cache = TRUE, message = FALSE, warning = FALSE}
nCores <- parallel::detectCores()

set.seed(888)

# Non-Parametric Bootstrap
bootNonParametric2 <- bootnet(networkCLIN2,
                             nBoots = 1000,
                             nCores = nCores - 1)
# Case-Drop Bootstrap
bootCaseDrop2 <- bootnet(networkCLIN2,
                        nBoots = 1000,
                        nCores = nCores - 1,
                        type = "case",
                        statistics = c("Strength", "Closeness", "Betweenness"))
```

```{r, message = FALSE, warning = FALSE}
# Plot Edge Accuracy Non-Parametric Bootstrap
plot(bootNonParametric2,
     order = "sample",
     labels = FALSE)

# Plot Edge Differences Non-Parametric Bootstrap
plot(bootNonParametric2,
     plot = "difference",
     onlyNonZero= TRUE,
     order = "sample")

# Plot Centrality Accuracy Case-Drop Bootstrap
plot(bootCaseDrop2,
     statistics = c("Strength", "Closeness", "Betweenness"))

# CS-Coefficient
corStability(bootCaseDrop2)

# Difference Test
plot(bootNonParametric2, "strength")
```

```{r, cache = TRUE, message = FALSE, warning = FALSE}
nCores <- parallel::detectCores()

set.seed(777)

# Non-Parametric Bootstrap
bootNonParametricNotCLIN2 <- bootnet(networkNotCLIN2,
                             nBoots = 1000,
                             nCores = nCores - 1)
# Case-Drop Bootstrap
bootCaseDropNotCLIN2 <- bootnet(networkNotCLIN2,
                        nBoots = 1000,
                        nCores = nCores - 1,
                        type = "case",
                        statistics = c("Strength", "Closeness", "Betweenness"))
```

```{r, message = FALSE, warning = FALSE}
# Plot Edge Accuracy Non-Parametric Bootstrap
plot(bootNonParametricNotCLIN2,
     order = "sample",
     labels = FALSE)

# Plot Edge Differences Non-Parametric Bootstrap
plot(bootNonParametricNotCLIN2,
     plot = "difference",
     onlyNonZero= TRUE,
     order = "sample")

# Plot Centrality Accuracy Case-Drop Bootstrap
plot(bootCaseDropNotCLIN2,
     statistics = c("Strength", "Closeness", "Betweenness"))

# CS-Coefficient
corStability(bootCaseDropNotCLIN2)

# Difference Test
plot(bootCaseDropNotCLIN2, "strength")
```

#### Mixed Graphical Model (MGM)

##### Predict all specializations

```{r}
# Create data frame excluding student IDs
dataMGM <- data_wide[,-2]

# Recode specialization to numeric 
dataMGM %<>% 
  mutate(specialization = as.numeric(as.factor(Specialization))) %>%
    select(-Specialization)

# Create sample for prediction
set.seed(65)
trainidx <- sample(seq_len(nrow(dataMGM)),
                   size = floor(.8 * nrow(dataMGM)))
```

```{r}
# Fitting a MGM with the mgm-package
set.seed(1)
fit_obj <- mgm(data = na.omit(dataMGM[trainidx,]), # na.omit excludes N = 32 
                type = c(rep("g", 5),"c"),
                level = c(rep(1, 5), 7),
                ruleReg = "OR",
                k = 2,
                binarySign = TRUE)
```

```{r}
# Node Predictability estimation
p_obj <- predict(fit_obj, na.omit(dataMGM[-trainidx,]),
                  errorCat = c("CC","; NCC","CCmarg"),
                  errorCon = c("R2"))

p_obj$errors

p_obj$predicted[,6] %>%
  table()

dataMGM[-trainidx,] %>%
  na.omit() %>%
    select(specialization) %>%
      table() %>%
        proportions()
```

```{r}
error_list <- list() # List for ring-segments
for(i in 1:5)
  error_list[[i]] <- p_obj$errors[i,2]
  beyondmarg <- p_obj$errors[6,3]-p_obj$errors[6,5]
  error_list[[6]] <- c(p_obj$errors[6,5],beyondmarg)

color_list <- list() # List for Colors

for(i in 1:5)
  color_list[[i]] <- "#90B4D4"
  color_list[[6]] <- c("#ffa500", "#ff4300")
```

```{r, fig.width=12, fig.height=12}
set.seed(10)
qgraph(fit_obj$pairwise$wadj,
       pie = error_list,
       layout="circle",
       theme = "colorblind",
       labels = colnames(dataMGM),
       pieColor = color_list,
       label.cex = .9,
       curveAll = TRUE,
       curveDefault = .6,
       cut = 0,
       labels = colnames(dataMGM))
```

##### Predict clinical vs non-clinical

```{r}
# Create data frame excluding student IDs
dataMGM <- data_wide[,-2]

# Recode specialization to numeric 
dataMGM %<>% 
  mutate(Clinical = ifelse(Specialization %in% "Spec Klinische Psychologie", 1, 0)) %>%
    select(-Specialization)

# Create sample for prediction
set.seed(25)
trainidx <- sample((seq_len(nrow(dataMGM))), size = floor(.8 * nrow(dataMGM)))
```

```{r}
# Fitting a MGM with the mgm-package
set.seed(1)
fit_obj <- mgm(data = na.omit(dataMGM[trainidx,]), # na.omit excludes N = 32 
                type = c(rep("g", 5),"c"),
                level = c(rep(1, 5), 2),
                ruleReg = "OR",
                k = 2,
                binarySign = TRUE)
```

```{r}
# Node Predictability estimation
p_obj <- predict(fit_obj, na.omit(dataMGM[-trainidx,]),
                  errorCat = c("CC","; NCC","CCmarg"),
                  errorCon = c("R2"))

p_obj$errors

p_obj$predicted[,6] %>%
  table()

dataMGM[-trainidx,] %>%
  na.omit() %>%
    select(Clinical) %>%
      table() %>%
        proportions()
```

```{r}
error_list <- list() # List for ring-segments
for(i in 1:5)
  error_list[[i]] <- p_obj$errors[i,2]
  beyondmarg <- p_obj$errors[6,3]-p_obj$errors[6,5]
  error_list[[6]] <- c(p_obj$errors[6,5],beyondmarg)

color_list <- list() # List for Colors

for(i in 1:5)
  color_list[[i]] <- "#90B4D4"
  color_list[[6]] <- c("#ffa500", "#ff4300")
```

```{r, fig.width=12, fig.height=12}
set.seed(10)
qgraph(fit_obj$pairwise$wadj,
       pie = error_list,
       layout="circle",
       theme = "colorblind",
       labels = colnames(dataMGM),
       pieColor = color_list,
       label.cex = .9,
       curveAll = TRUE,
       curveDefault = .6,
       cut = 0,
       labels = colnames(dataMGM))
```

##### Prediction Attempt - Train/Test Set

```{r}
# Create data frame excluding student IDs
dataMGM <- data_wide[,-2]

# Recode specialization to numeric 
dataMGM %<>% 
  mutate(Clinical = ifelse(Specialization %in% c("Spec Klinische Psychologie",
                                                "Spec Klinische Ontwikkelingsps",
                                                "Spec Klinische Neuropsych.",
                                                "Spec Brein en Cognitie"), 1, 0)) %>%
    select(-Specialization)

# Create sample for prediction
set.seed(2)
trainidx <- sample((seq_len(nrow(dataMGM))), size = floor(.8 * nrow(dataMGM)))
```

```{r}
# Fitting a MGM with the mgm-package
set.seed(1)
fit_obj <- mgm(data = na.omit(dataMGM[trainidx,]), # na.omit excludes N = 32 
                type = c(rep("g", 5),"c"),
                level = c(rep(1, 5), 2),
                ruleReg = "OR",
                k = 2,
                binarySign = TRUE)
```

```{r}
# Node Predictability estimation
p_obj <- predict(fit_obj, na.omit(dataMGM[-trainidx,]),
                  errorCat = c("CC","; NCC","CCmarg"),
                  errorCon = c("R2"))

p_obj$predicted[,6] %>%
  table()

dataMGM[-trainidx,] %>%
  na.omit() %>%
    select(Clinical) %>%
      table() %>%
        proportions()
```

```{r}
error_list <- list() # List for ring-segments
for(i in 1:5)
  error_list[[i]] <- p_obj$errors[i,2]
  beyondmarg <- p_obj$errors[6,3]-p_obj$errors[6,5]
  error_list[[6]] <- c(p_obj$errors[6,5],beyondmarg)

color_list <- list() # List for Colors

for(i in 1:5)
  color_list[[i]] <- "#90B4D4"
  color_list[[6]] <- c("#ffa500", "#ff4300")
```

```{r, fig.width=12, fig.height=12}
set.seed(10)
qgraph(fit_obj$pairwise$wadj,
       pie = error_list,
       layout="circle",
       theme = "colorblind",
       labels = colnames(dataMGM),
       pieColor = color_list,
       label.cex = .9,
       curveAll = TRUE,
       curveDefault = .6,
       cut = 0,
       labels = colnames(dataMGM))
```

##### Prediction Attempt - Cohort

```{r}
dataMGM <- full_tidy_data %>%
  select(specialisation, grade, course_id, student_id, cohort) %>%
  # Retain only first year courses (shared programme)
  filter(course_id %in% selected_courses) %>% 
    group_by(student_id) %>%
      pivot_wider(.,
                  names_from = course_id,
                  values_from = grade) %>%
          ungroup()

# Assigning uniform column names for the courses
colnames(dataMGM) <-
  c("Specialization",
    "ID",
    "Cohort",           # Include Cohort here
    "IPC",              #7201602PXY
    "RMS",              #7201604PXY
    "DEV",              #7201605PXY
    "SWO",              #7201607PXY
    "CB")               #7201610PXY

table(full_tidy_data$cohort)
table(dataMGM$Cohort)
```

```{r}
# Create training sample (1718 + 1819)
dataMGM1719 <- filter(dataMGM, Cohort != "1920")
dataMGM1920 <- filter(dataMGM, Cohort %in% "1920")

# Recode specialization to numeric 
dataMGM1719 %<>% 
  mutate(Clinical = ifelse(Specialization %in% "Spec Klinische Psychologie", 1, 0)) %>%
    select(-Specialization, -ID, -Cohort)

set.seed(111)
dataMGM1920 %<>%
  select(-Specialization, -ID, -Cohort) %>%
    mutate(Clinical = sample(0:1, nrow(dataMGM1920), replace = TRUE))
```

```{r}
# Fitting a MGM with the mgm-package
set.seed(1)
fit_obj <- mgm(data = na.omit(dataMGM1719), # na.omit excludes N = 32 
                type = c(rep("g", 5),"c"),
                level = c(rep(1, 5), 2),
                ruleReg = "OR",
                k = 2,
                binarySign = TRUE)
```

```{r}
# Node Predictability estimation
p_obj <- predict(fit_obj, 
                 na.omit(dataMGM1920),
                 errorCat = c("CC","; NCC","CCmarg"),
                 errorCon = c("R2"))

p_obj$errors

p_obj$predicted[,6] %>%
  table()

dataMGM1920 %>%
  na.omit() %>%
    select(Clinical) %>%
      table() %>%
        proportions()
```

```{r}
error_list <- list() # List for ring-segments
for(i in 1:5)
  error_list[[i]] <- p_obj$errors[i,2]
  beyondmarg <- p_obj$errors[6,3]-p_obj$errors[6,5]
  error_list[[6]] <- c(p_obj$errors[6,5],beyondmarg)

color_list <- list() # List for Colors

for(i in 1:5)
  color_list[[i]] <- "#90B4D4"
  color_list[[6]] <- c("#ffa500", "#ff4300")
```

```{r, fig.width=12, fig.height=12}
set.seed(10)
qgraph(fit_obj$pairwise$wadj,
       pie = error_list,
       layout="circle",
       theme = "colorblind",
       labels = colnames(dataMGM1920),
       pieColor = color_list,
       label.cex = .9,
       curveAll = TRUE,
       curveDefault = .6,
       cut = 0,
       labels = colnames(dataMGM1920))
```

