---
title: "Untitled"
author: "Wesley Kammeraad"
date: "20-1-2021"
output: html_document
---

```{r, echo = FALSE, message = FALSE, warning = FALSE}
installed_packages <- installed.packages()[, "Package"]

## Install Missing Required Packages
if (!"tidyverse" %in% installed_packages) {
  install.packages("tidyverse")
  }
if (!"qgraph" %in% installed_packages) {
  install.packages("qgraph")
}
if (!"dplyr" %in% installed_packages) {
  install.packages("dplyr")
}
if (!"psych" %in% installed_packages) {
  install.packages("magrittr")
}
if (!"bootnet" %in% installed_packages) {
  install.packages("bootnet")
}
if (!"mgm" %in% installed_packages) {
  install.packages("mgm")
}

## Load packages
library("tidyverse")
library("qgraph")
library("dplyr")
library("magrittr")
library("bootnet")
library("mgm")
```

## %in%; Study Guide Number

```{r}
data1718 <- read.csv("data/anonymized_cohort1718.csv",
                     header = TRUE,
                     sep = ";",
                     dec = ",",
                     na.strings = c("AVV", "-"))

data1819 <- read.csv("data/anonymized_cohort1819.csv",
                     header = TRUE,
                     sep = ";",
                     dec = ",",
                     na.strings = c("AVV", "-"))

data1920 <- read.csv("data/anonymized_cohort1920.csv",
                     header = TRUE,
                     sep = ";",
                     dec = ",",
                     na.strings = c("AVV", "-"))
```

```{r}
colnames(data1718) <- c("Track", "Subject", "Country", "Specialization", "Grade", "Course", "CourseDesc", "Study Guide", "ID")

colnames(data1819) <- c("Track", "Subject", "Country", "Specialization", "Grade", "Course", "CourseDesc", "Study Guide", "Rank", "ID")

colnames(data1920) <- c("Track", "Subject", "Country", "Grade", "Course", "CourseDesc", "Study Guide", "Rank", "ID")  
```

```{r}
# Changing grades from character to numeric data
data1718$Grade <- gsub(",", ".", data1718$Grade)
data1718$Grade <- as.numeric(data1718$Grade)

data1819$Grade <- gsub(",", ".", data1819$Grade)
data1819$Grade <- as.numeric(data1819$Grade)
```

```{r}
data1718 %>%
  select(CourseDesc, `Study Guide`) %>%
    unique()

data1718 %>%
  select(`Study Guide`) %>%
    unique()


data1718 %>%
  select(CourseDesc) %>%
    unique()
```
Shared Program 1718

> Introductory Psychology and Cognition	7201602PXY			
> Research Methods and Statistics	7201604PXY			
> Developmental Psychology	7201605PXY			
> Social and Work & Organisational Psychology	7201607PXY			
> Clinical Psychology & Brain and Cognition	7201622PXY

Shared Program 1819

> Introductory Psychology and Cognition	7201602PXY			 
> Research Methods and Statistics	7201604PXY			
> Developmental Psychology	7201605PXY			
> Social and Work & Organisational Psychology	7201607PXY			
> Clinical Psychology & Brain and Cognition	7201622PXY

```{r}
# Retain Shared Program Courses
courses1718 <- 
  data1718 %>%
    select(CourseDesc, `Study Guide`) %>%
      unique() %>%
        filter(CourseDesc %in% c("Introductory Psychology and Cognition",
                               "Research Methods and Statistics",
                               "Developmental Psychology",
                               "Social and Work & Organisational Psychology",
                               "Clinical and Biological Psychology"))

courses1819 <- 
  data1819 %>%
    select(CourseDesc, `Study Guide`) %>%
      unique() %>%
        filter(CourseDesc %in% c("Introductory Psychology and Cognition",
                               "Research Methods and Statistics",
                               "Developmental Psychology",
                               "Social and Work & Organisational Psychology",
                               "Clinical Psychology & Brain and Cognition"))
```

```{r}
data1718 <- data1718[data1718$`Study Guide` %in% courses1718$`Study Guide`,]
data1819 <- data1819[data1819$`Study Guide` %in% courses1819$`Study Guide`,]
```

```{r}
data1718 %<>%
  select(Specialization, Grade, Course, ID) %>%
  pivot_wider(.,
              names_from = Course,
              values_from = Grade)

data1819 %<>%
  select(Specialization, Grade, Course, ID) %>%
  pivot_wider(.,
              names_from = Course,
              values_from = Grade)
```

```{r}
# Assigning uniform column names for the courses
colnames(data1718) <-
  colnames(data1819) <-
  c("Specialization",
    "ID",
    "IP&C",
    "RM&S",
    "DevPsy",
    "S&WO",
    "C&B")

data1718
data1819
```

```{r}
data <- rbind(data1718, data1819)
data
```

```{r}
data %>%
  distinct(ID, Specialization) %>%
    count(Specialization)
```

```{r}
# Removing all students that have not picked a specialization
dataSpec <- data[data$Specialization != 0,]
dataSpec
```

```{r}
dataSpec %>%
  distinct(ID, Specialization) %>%
    count(Specialization)
```

```{r}
network <- estimateNetwork(dataSpec[,-c(1:2)],
                           default = "pcor",
                           corMethod = "spearman")

qgraph(network$graph,
       layout = "spring",
       theme = "colorblind")
```

-------------------------------------------------------------------------------------------

## Specialization Dummy Code

### Polychoric Correlations; Spearman

```{r}
# Create Dummy Variable for "CLIN" Specialization: Yes/No
dataCLIN <- 
  dataSpec %>%
      mutate(CLIN = ifelse(dataSpec$Specialization == "Spec Klinische Psychologie", 1, 0))


networkCLIN <- estimateNetwork(dataCLIN[,-c(1:2)],
                               default = "pcor",
                               corMethod = "spearman")

qgraph(networkCLIN$graph,
       layout = "spring",
       theme = "colorblind",
       title = paste("N =", nrow(dataCLIN)))
```

```{r}
# Create Dummy Variable for "WOP" Specialization: Yes/No
dataWOP <- 
  dataSpec %>%
      mutate(WOP = ifelse(dataSpec$Specialization == "Spec Arbeids- en Org. Psych.", 1, 0))


networkWOP <- estimateNetwork(dataWOP[,-c(1:2)],
                               default = "pcor",
                               corMethod = "spearman")

qgraph(networkWOP$graph,
       layout = "spring",
       theme = "colorblind",
       title = paste("N =", nrow(dataWOP)))
```

```{r}
# Create Dummy Variable for "SOC" Specialization: Yes/No
dataSOC <- 
  dataSpec %>%
      mutate(SOC = ifelse(dataSpec$Specialization == "Spec Sociale Psychologie", 1, 0))


networkSOC <- estimateNetwork(dataSOC[,-c(1:2)],
                               default = "pcor",
                               corMethod = "spearman")

qgraph(networkSOC$graph,
       layout = "spring",
       theme = "colorblind",
       title = paste("N =", nrow(dataSOC)))
```

```{r}
# Create Dummy Variable for "BC" Specialization: Yes/No
dataBC <- 
  dataSpec %>%
      mutate(BC = ifelse(dataSpec$Specialization == "Spec Brein en Cognitie", 1, 0))


networkBC <- estimateNetwork(dataBC[,-c(1:2)],
                               default = "pcor",
                               corMethod = "spearman")

qgraph(networkBC$graph,
       layout = "spring",
       theme = "colorblind",
       title = paste("N =", nrow(dataBC)))
```

```{r}
# Create Dummy Variable for "DEV" Specialization: Yes/No
dataDEV <- 
  dataSpec %>%
      mutate(DEV = ifelse(dataSpec$Specialization == "Spec Klinische Ontwikkelingsps", 1, 0))


networkDEV <- estimateNetwork(dataDEV[,-c(1:2)],
                               default = "pcor",
                               corMethod = "spearman")

qgraph(networkDEV$graph,
       layout = "spring",
       theme = "colorblind",
       title = paste("N =", nrow(dataDEV)))
```

```{r}
# Create Dummy Variable for "PML" Specialization: Yes/No
dataPML <- 
  dataSpec %>%
      mutate(PML = ifelse(dataSpec$Specialization == "Spec Psych. Methodenleer", 1, 0))


networkPML <- estimateNetwork(dataPML[,-c(1:2)],
                               default = "pcor",
                               corMethod = "spearman")

qgraph(networkPML$graph,
       layout = "spring",
       theme = "colorblind",
       title = paste("N =", nrow(dataPML)))
```

### EBICglasso; Spearman

```{r}
# Create Dummy Variable for "CLIN" Specialization: Yes/No
dataCLIN <- 
  dataSpec %>%
      mutate(CLIN = ifelse(dataSpec$Specialization == "Spec Klinische Psychologie", 1, 0))


networkCLIN <- estimateNetwork(dataCLIN[,-c(1:2)],
                               default = "EBICglasso",
                               corMethod = "spearman")

qgraph(networkCLIN$graph,
       layout = "spring",
       theme = "colorblind",
       title = paste("N =", nrow(dataCLIN)))
```

```{r}
# Create Dummy Variable for "WOP" Specialization: Yes/No
dataWOP <- 
  dataSpec %>%
      mutate(WOP = ifelse(dataSpec$Specialization == "Spec Arbeids- en Org. Psych.", 1, 0))


networkWOP <- estimateNetwork(dataWOP[,-c(1:2)],
                               default = "EBICglasso",
                               corMethod = "spearman")

qgraph(networkWOP$graph,
       layout = "spring",
       theme = "colorblind",
       title = paste("N =", nrow(dataWOP)))
```

```{r}
# Create Dummy Variable for "SOC" Specialization: Yes/No
dataSOC <- 
  dataSpec %>%
      mutate(SOC = ifelse(dataSpec$Specialization == "Spec Sociale Psychologie", 1, 0))


networkSOC <- estimateNetwork(dataSOC[,-c(1:2)],
                               default = "EBICglasso",
                               corMethod = "spearman")

qgraph(networkSOC$graph,
       layout = "spring",
       theme = "colorblind",
       title = paste("N =", nrow(dataSOC)))
```

```{r}
# Create Dummy Variable for "BC" Specialization: Yes/No
dataBC <- 
  dataSpec %>%
      mutate(BC = ifelse(dataSpec$Specialization == "Spec Brein en Cognitie", 1, 0))


networkBC <- estimateNetwork(dataBC[,-c(1:2)],
                               default = "EBICglasso",
                               corMethod = "spearman")

qgraph(networkBC$graph,
       layout = "spring",
       theme = "colorblind",
       title = paste("N =", nrow(dataBC)))
```

```{r}
# Create Dummy Variable for "DEV" Specialization: Yes/No
dataDEV <- 
  dataSpec %>%
      mutate(DEV = ifelse(dataSpec$Specialization == "Spec Klinische Ontwikkelingsps", 1, 0))


networkDEV <- estimateNetwork(dataDEV[,-c(1:2)],
                               default = "EBICglasso",
                               corMethod = "spearman")

qgraph(networkDEV$graph,
       layout = "spring",
       theme = "colorblind",
       title = paste("N =", nrow(dataDEV)))
```

```{r}
# Create Dummy Variable for "PML" Specialization: Yes/No
dataPML <- 
  dataSpec %>%
      mutate(PML = ifelse(dataSpec$Specialization == "Spec Psych. Methodenleer", 1, 0))


networkPML <- estimateNetwork(dataPML[,-c(1:2)],
                               default = "EBICglasso",
                               corMethod = "spearman")

qgraph(networkPML$graph,
       layout = "spring",
       theme = "colorblind",
       title = paste("N =", nrow(dataPML)))
```

### EBICglasso; cor_auto

```{r}
# Create Dummy Variable for "CLIN" Specialization: Yes/No
dataCLIN <- 
  dataSpec %>%
      mutate(CLIN = ifelse(dataSpec$Specialization == "Spec Klinische Psychologie", 1, 0))


networkCLIN <- estimateNetwork(dataCLIN[,-c(1:2)],
                               default = "EBICglasso",
                               corMethod = "cor_auto")

qgraph(networkCLIN$graph,
       layout = "spring",
       theme = "colorblind",
       title = paste("N =", nrow(dataCLIN)))
```

```{r}
# Create Dummy Variable for "WOP" Specialization: Yes/No
dataWOP <- 
  dataSpec %>%
      mutate(WOP = ifelse(dataSpec$Specialization == "Spec Arbeids- en Org. Psych.", 1, 0))


networkWOP <- estimateNetwork(dataWOP[,-c(1:2)],
                               default = "EBICglasso",
                               corMethod = "cor_auto")

qgraph(networkWOP$graph,
       layout = "spring",
       theme = "colorblind",
       title = paste("N =", nrow(dataWOP)))
```

```{r}
# Create Dummy Variable for "SOC" Specialization: Yes/No
dataSOC <- 
  dataSpec %>%
      mutate(SOC = ifelse(dataSpec$Specialization == "Spec Sociale Psychologie", 1, 0))


networkSOC <- estimateNetwork(dataSOC[,-c(1:2)],
                               default = "EBICglasso",
                               corMethod = "cor_auto")

qgraph(networkSOC$graph,
       layout = "spring",
       theme = "colorblind",
       title = paste("N =", nrow(dataSOC)))
```

```{r}
# Create Dummy Variable for "BC" Specialization: Yes/No
dataBC <- 
  dataSpec %>%
      mutate(BC = ifelse(dataSpec$Specialization == "Spec Brein en Cognitie", 1, 0))


networkBC <- estimateNetwork(dataBC[,-c(1:2)],
                               default = "EBICglasso",
                               corMethod = "cor_auto")

qgraph(networkBC$graph,
       layout = "spring",
       theme = "colorblind",
       title = paste("N =", nrow(dataBC)))
```

```{r}
# Create Dummy Variable for "DEV" Specialization: Yes/No
dataDEV <- 
  dataSpec %>%
      mutate(DEV = ifelse(dataSpec$Specialization == "Spec Klinische Ontwikkelingsps", 1, 0))


networkDEV <- estimateNetwork(dataDEV[,-c(1:2)],
                               default = "EBICglasso",
                               corMethod = "cor_auto")

qgraph(networkDEV$graph,
       layout = "spring",
       theme = "colorblind",
       title = paste("N =", nrow(dataDEV)))
```

```{r}
# Create Dummy Variable for "PML" Specialization: Yes/No
dataPML <- 
  dataSpec %>%
      mutate(PML = ifelse(dataSpec$Specialization == "Spec Psych. Methodenleer", 1, 0))


networkPML <- estimateNetwork(dataPML[,-c(1:2)],
                               default = "EBICglasso",
                               corMethod = "cor_auto")

qgraph(networkPML$graph,
       layout = "spring",
       theme = "colorblind",
       title = paste("N =", nrow(dataPML)))
```


-------------------------------------------------------------------------------------------

## Specialization Data Split

### Polychoric Correlations; Spearman

```{r}
# Estimate Network for All students who have specialized in Clinical
dataCLIN <- 
  dataSpec %>%
      filter(dataSpec$Specialization == "Spec Klinische Psychologie")


networkCLIN <- estimateNetwork(dataCLIN[,-c(1:2)],
                               default = "pcor",
                               corMethod = "spearman")

qgraph(networkCLIN$graph,
       layout = "spring",
       theme = "colorblind",
       title = paste("N =", nrow(dataCLIN))) # N = 45
```

```{r}
# Estimate Network for All students who have specialized in Work & Organisational 
dataWOP <- 
  dataSpec %>%
      filter(dataSpec$Specialization == "Spec Arbeids- en Org. Psych.")


networkWOP <- estimateNetwork(dataWOP[,-c(1:2)],
                               default = "pcor",
                               corMethod = "spearman")

qgraph(networkWOP$graph,
       layout = "spring",
       theme = "colorblind",
       title = paste("N =", nrow(dataWOP))) # N = 22
```

```{r, eval = FALSE}
# Estimate Network for All students who have specialized in Social
dataSOC <- 
  dataSpec %>%
    filter(dataSpec$Specialization == "Spec Sociale Psychologie")

networkSOC <- estimateNetwork(dataSOC[,-c(1:2)],
                               default = "pcor",
                               corMethod = "spearman")

qgraph(networkSOC$graph,
       layout = "spring",
       theme = "colorblind",
       title = paste("N =", nrow(dataSOC))) # N = 15
```

```{r, eval = FALSE}
# Estimate Network for All students who have specialized in Brain & Cognition
dataBC <- 
  dataSpec %>%
      filter(dataSpec$Specialization == "Spec Brein en Cognitie")


networkBC <- estimateNetwork(dataBC[,-c(1:2)],
                               default = "pcor",
                               corMethod = "spearman")

qgraph(networkBC$graph,
       layout = "spring",
       theme = "colorblind",
       title = paste("N =", nrow(dataBC))) # N = 11
```

```{r}
# Estimate Network for All students who have specialized in Developmental
dataDEV <- 
  dataSpec %>%
      filter(dataSpec$Specialization == "Spec Klinische Ontwikkelingsps")

networkDEV <- estimateNetwork(dataDEV[,-c(1:2)],
                               default = "pcor",
                               corMethod = "spearman")

qgraph(networkDEV$graph,
       layout = "spring",
       theme = "colorblind",
       title = paste("N =", nrow(dataDEV))) # N = 14
```

```{r, eval = FALSE}
# Estimate Network for All students who have specialized in Methods
dataPML <- 
  dataSpec %>%
      filter(dataSpec$Specialization == "Spec Psych. Methodenleer")

networkPML <- estimateNetwork(dataPML[,-c(1:2)],
                               default = "pcor",
                               corMethod = "spearman")

qgraph(networkPML$graph,
       layout = "spring",
       theme = "colorblind",
       title = paste("N =", nrow(dataPML))) # N = 4
```

### EBICglasso; Spearman

```{r}
# Estimate Network for All students who have specialized in Clinical
dataCLIN <- 
  dataSpec %>%
      filter(dataSpec$Specialization == "Spec Klinische Psychologie")


networkCLIN <- estimateNetwork(dataCLIN[,-c(1:2)],
                               default = "EBICglasso",
                               corMethod = "spearman")

qgraph(networkCLIN$graph,
       layout = "spring",
       theme = "colorblind",
       title = paste("N =", nrow(dataCLIN))) # N = 45
```

```{r}
# Estimate Network for All students who have specialized in Work & Organisational 
dataWOP <- 
  dataSpec %>%
      filter(dataSpec$Specialization == "Spec Arbeids- en Org. Psych.")


networkWOP <- estimateNetwork(dataWOP[,-c(1:2)],
                               default = "EBICglasso",
                               corMethod = "spearman")

qgraph(networkWOP$graph,
       layout = "spring",
       theme = "colorblind",
       title = paste("N =", nrow(dataWOP))) # N = 22
```

```{r, eval = FALSE}
# Estimate Network for All students who have specialized in Social
dataSOC <- 
  dataSpec %>%
    filter(dataSpec$Specialization == "Spec Sociale Psychologie")

networkSOC <- estimateNetwork(dataSOC[,-c(1:2)],
                               default = "EBICglasso",
                               corMethod = "spearman")

qgraph(networkSOC$graph,
       layout = "spring",
       theme = "colorblind",
       title = paste("N =", nrow(dataSOC))) # N = 15
```

```{r, eval = FALSE}
# Estimate Network for All students who have specialized in Brain & Cognition
dataBC <- 
  dataSpec %>%
      filter(dataSpec$Specialization == "Spec Brein en Cognitie")


networkBC <- estimateNetwork(dataBC[,-c(1:2)],
                               default = "EBICglasso",
                               corMethod = "spearman")

qgraph(networkBC$graph,
       layout = "spring",
       theme = "colorblind",
       title = paste("N =", nrow(dataBC))) # N = 11
```

```{r}
# Estimate Network for All students who have specialized in Developmental
dataDEV <- 
  dataSpec %>%
      filter(dataSpec$Specialization == "Spec Klinische Ontwikkelingsps")

networkDEV <- estimateNetwork(dataDEV[,-c(1:2)],
                               default = "EBICglasso",
                               corMethod = "spearman")

qgraph(networkDEV$graph,
       layout = "spring",
       theme = "colorblind",
       title = paste("N =", nrow(dataDEV))) # N = 14
```

```{r, eval = FALSE}
# Estimate Network for All students who have specialized in Methods
dataPML <- 
  dataSpec %>%
      filter(dataSpec$Specialization == "Spec Psych. Methodenleer")

networkPML <- estimateNetwork(dataPML[,-c(1:2)],
                               default = "EBICglasso",
                               corMethod = "spearman")

qgraph(networkPML$graph,
       layout = "spring",
       theme = "colorblind",
       title = paste("N =", nrow(dataPML))) # N = 4
```

### EBICglasso; cor_auto

```{r}
# Estimate Network for All students who have specialized in Clinical
dataCLIN <- 
  dataSpec %>%
      filter(dataSpec$Specialization == "Spec Klinische Psychologie")


networkCLIN <- estimateNetwork(dataCLIN[,-c(1:2)],
                               default = "EBICglasso",
                               corMethod = "cor_auto")

qgraph(networkCLIN$graph,
       layout = "spring",
       theme = "colorblind",
       title = paste("N =", nrow(dataCLIN))) # N = 45
```

```{r}
# Estimate Network for All students who have specialized in Work & Organisational 
dataWOP <- 
  dataSpec %>%
      filter(dataSpec$Specialization == "Spec Arbeids- en Org. Psych.")


networkWOP <- estimateNetwork(dataWOP[,-c(1:2)],
                               default = "EBICglasso",
                               corMethod = "cor_auto")

qgraph(networkWOP$graph,
       layout = "spring",
       theme = "colorblind",
       title = paste("N =", nrow(dataWOP))) # N = 22
```

```{r, eval = FALSE}
# Estimate Network for All students who have specialized in Social
dataSOC <- 
  dataSpec %>%
    filter(dataSpec$Specialization == "Spec Sociale Psychologie")

networkSOC <- estimateNetwork(dataSOC[,-c(1:2)],
                               default = "EBICglasso",
                               corMethod = "cor_auto")

qgraph(networkSOC$graph,
       layout = "spring",
       theme = "colorblind",
       title = paste("N =", nrow(dataSOC))) # N = 15
```

```{r, eval = FALSE}
# Estimate Network for All students who have specialized in Brain & Cognition
dataBC <- 
  dataSpec %>%
      filter(dataSpec$Specialization == "Spec Brein en Cognitie")


networkBC <- estimateNetwork(dataBC[,-c(1:2)],
                               default = "EBICglasso",
                               corMethod = "cor_auto")

qgraph(networkBC$graph,
       layout = "spring",
       theme = "colorblind",
       title = paste("N =", nrow(dataBC))) # N = 11
```

```{r}
# Estimate Network for All students who have specialized in Developmental
dataDEV <- 
  dataSpec %>%
      filter(dataSpec$Specialization == "Spec Klinische Ontwikkelingsps")

networkDEV <- estimateNetwork(dataDEV[,-c(1:2)],
                               default = "EBICglasso",
                               corMethod = "cor_auto")

qgraph(networkDEV$graph,
       layout = "spring",
       theme = "colorblind",
       title = paste("N =", nrow(dataDEV))) # N = 14
```

```{r, eval = FALSE}
# Estimate Network for All students who have specialized in Methods
dataPML <- 
  dataSpec %>%
      filter(dataSpec$Specialization == "Spec Psych. Methodenleer")

networkPML <- estimateNetwork(dataPML[,-c(1:2)],
                               default = "EBICglasso",
                               corMethod = "cor_auto")

qgraph(networkPML$graph,
       layout = "spring",
       theme = "colorblind",
       title = paste("N =", nrow(dataPML))) # N = 4
```

***

## Node Predictability

```{r}
dataSpec$Specialization <- factor(dataSpec$Specialization, 
                                  ordered = FALSE)

dataSpec$Specialization <- recode(dataSpec$Specialization,
                                  "Spec Klinische Psychologie" = "1",
                                  "Spec Arbeids- en Org. Psych." = "2",
                                  "Spec Sociale Psychologie" = "3",
                                  "Spec Klinische Ontwikkelingsps" = "4",
                                  "Spec Brein en Cognitie" = "5",
                                  "Spec Psych. Methodenleer" = "6")

dataSpec[,c(3:7)] <- sapply(dataSpec[,c(3:7)], function(x){round(x, 1)})
dataSpec[,c(3:7)] <- sapply(dataSpec[,c(3:7)], as.numeric)
dataSpec
```

```{r}
type <- c("c", "g", "g", "g", "g", "g")
level <- c(6, 1, 1, 1, 1, 1)

mgm(data = na.omit(dataSpec)[,-2],
    type = type,
    level = level,
    k = 3)
```

xxxx